---
title: "Dynmatic predtion model in parkinson disease"
author: "Yan Li"
header-includes:
- \usepackage{floatrow}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \floatsetup[table]{capposition=top}
- \floatplacement{table}{H}
output:
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
        scroll_highlight: yes
      before: null
      after: null
    toolbar:
      position: fixed
    edit : null
    download: [["book.pdf", "Download PDF version"]]
    search: yes
    fontsettings:
      theme: white
      family: sans
      size: 2
    sharing:
      facebook: yes
      github: no
      twitter: yes
      linkedin: no
      weibo: no
      instapaper: no
      vk: no
      whatsapp: no
      all: ['facebook', 'twitter', 'linkedin', 'weibo', 'instapaper']
    info: yes
  bookdown::html_book:
    css: toc.css
  bookdown::pdf_book:
    keep_tex: yes
documentclass: book
bibliography: my5.bib
fontsize: 14pt
citation_package: none
csl: cardiff-university-vancouver.csl
nocite: '@*'
---
# Systematic review of prognostic models in Parkinson’s disease

# PD cox regression

## Abstract

Background

When patients with parkinsonism lose their ability to take care of themselves, and insufficient care is available at home, they will need long-term care in a nursing home or other institution (institutionalisation). The risk factors for institutionalisation are unclear and there are few published studies. We aimed to identify the incidence of, and the risk factors for, institutionalization in Parkinson's disease (PD) and atypical parkinsonism.

Methods

We analysed data from a prospective population-based incidence cohort of parkinsonism in North East Scotland (the PINE study). 556 participants(Parkinson's disease, N=200; atypical parkinsonism, N=98; controls, N=258), recruited between 2002 and 2009, were prospectively followed life-long to May 2020 with data collection on place of residence. We calculated the incidence of institutionalisation and used Cox regression to determine baseline predictors of institutionalisation. 

Results

Over a median follow-up time 9.3 years in Parkinson’s disease and 4.4 years in atypical parkinsonism, 70 with PD, 53 with atypical parkinsonism, and 43 controls became institutionalised. The incidence rates of institutionalisation per 100 person-years in PD, atypical parkinsonism and controls were 5.2, 20.8, and 1.8 respectively. The Kaplan-Meier median time to institutionalization was 11.8 years in PD,  3.5 years in atypical parkinsonism. Multivariable Cox regression showed that atypical parkinsonism [HR versus PD (95% CI)=3.02 (1.84,4.96)], increasing age [HR for 10-year increase (95% CI)=1.89 (1.43,2.51)], poorer cognition [HR for MMSE<24 versus MMSE>27 (95% CI)=2.60 (1.43, 4.72)] were independently associated with higher hazards of institutionalization. Sex, Charlson co-morbidity index, smoking history, living alone, UPDRS part 3 (motor) score, and Schwab and England scale, were not associated with institutionalization. 

Conclusion

Institutionalisation is more frequent in parkinsonism, particularly atypical parkinsonism than in controls. Atypical parkinsonism, older age and poorer cognition were independent predictors. 







## Introduction

Parkinson’s disease (PD) and atypical parkinsonism are neurodegenerative disorders. The cause of PD is unknown, and both the genetic and environmental factors affect it. Moreover, it is difficult to diagnose PD and there are no treatments that can slow the underlying neurodegeneration, although there are effective symptomatic treatments, principally levodopa and other dopaminergic drugs. As the disease progresses, features which are resistant to dopaminergic therapies often become troublesome, including both motor (dysphagia, postural instability, freezing of gait and falls), and non-motor features (urinary incontinence, orthostatic hypotension, and dementia). These symptoms reduce patients’ quality of life and may lead to the loss of the ability to live independently. Care may be provided at home by relatives or professional carers, but this is not always available or sufficient to maintain living at home. Hence, many patients will require long-term care in an institution, such as a nursing home [@Lancet2015]. 

Only a few studies have investigated risk factors for institutionalisation. The first study was a retrospective case-control study with only 11 PD patients and found hallucinations was the risk factor for nursing home placement  [@pred1993]. Then, a four-year prospective population-based prevalence study was carried out in western Norway [@pred2000], which followed 178 community-dwelling subjects in PD from baseline to nursing home admission, and found the highest risk were hallucinations, cognitive impairment and old age. A nationwide retrospective cohort study followed 469055 elderly PD patients in the US from 2002 to December 2005 and found women and black individuals with PD were more likely to enter a nursing facility residence compared with men and white individuals [@pred2015]. However, no previous inception studies have reported the incidence of institutionalisation in PD or its risk factors. To our knowledge, no previous studies have investigated institutionalisation in atypical parkinsonian syndromes other than dementia with Lewy bodies. 

In the late stage of PD, the economic burden for patients is the cost of home care instead of drug therapy [@USburden;@UKburden]. And a study in Norway showed that if clinicians can help to delay PD patients being admitted into nursing home for 3 months, it can save approximating drug therapy yearly cost for patients [@Norway]. Therefore, finding the risk factors for institutionalization can help PD patients and health care providers reduce the economic burden by extending their independent living ability. The results from this study will enable clinicians to provide more personalized and timely management strategies to help PD patients. Using data from Parkinsonism incidence in North-East Scotland (PINE) study, a population-based incidence cohort, we aimed to: 

* Identify the incidence of institutionalization in PD and atypical parkinsonism and compare with age and sex matched controls. 
* Identify the risk factors for institutionalization in PD and atypical parkinsonism. 



## Methods

**Study population and design**

The Parkinsonism incidence in North-East Scotland (PINE) study is a prospective and population-based incidence cohort that sought to identify and follow all newly diagnosed patients with degenerative or vascular parkinsonism in Aberdeen, Scotland between 2002-4 and 2006-9 [@PINE]. All incidence patients were invited to undergo lifelong in-person annual follow-up. An age-sex matched control group was also recruited. At baseline, and at each annual visit, data are collected on demographics, clinical history, and non-motor symptoms, and a structured clinical examination and several questionnaires.  Data on key outcomes are collected, includingsurvival, dependency, dementia, and institutionalization up to 2020 [@Field2016]. In this study, there are 201 PD patients, 114 atypical parkinsonism patients (progressive supranuclear palsy, corticobasal degeneration, multiple system atrophy, and vascular parkinsonism) and 260 control individuals. The baseline visit time was from November 2002 to June 2010, and the last follow up time was April 2020. 185 individuals entered (45 in control group) an institution, 420 deaths occurred (142 in control group) and 6 were lost to follow up (3 in PD and 3 in control group).

**Potential prognostic factors**

Parkinson’s rating scales are used to assess patients’ symptoms of the condition and evaluate treatment. Clinicians assign a value depending on a patients’ particular feature information. In our study, there are three rating scales and one cognitive function measurement:

* The Unified Parkinson’s Disease Rating Scale (UPDRS): it has four parts, the first three sections evaluate the degree of disability and the fourth section evaluates any complications of treatment. And the UPDRS part 3 is motor examination and higher score indicates increase severity (0-4 rating scale). The UPDRS is often used with the Hoehn and Yahr Scale and the Schwab and England Activities of Daily Living (ADL) Scale [@Ratingscales]. 

* The Hoehn and Yahr Scale: it is used to measure the progress of Parkinson’s symptoms and the level of disability. It includes stages 0 to 5, where stage 0 means no signs of disease and stage 5 means the patient needs a wheelchair or is bedridden. 

* The Schwab and England ADL Scale: it is used to measure patient independence of daily activities through a percentage figure, where the higher the percentage the higher the level of independence. Patients’ comorbid conditions and cognitive function were also measured. Charlson co-morbidity index categorizes patients’ comorbid conditions by assigned a weight from 1 to 6 to each condition and it includes 16 conditions [@Charlson]. *

* The Mini Mental State Examination (MMSE) is used to assess the severity of cognitive impairment and the score of MMSE ranges from zero to 30 [@MMSE1]. And, we divided the MMSE into three categories: less than 24 (probable dementia), from 24 to 27 (probable cognitive impairment) and from 28 to 30 (normal cognitive function). The MMSE score less than 24 is commonly used to indicate dementia and the 27/28 threshold reflects the mean MMSE score in normal healthy adults of 27.8 [@MMSE2].


The annual clinical examination includes: (i) parkinsonian impairment (UPDRS part3 (motor) score); (ii) disease stage (Hoehn and Yahr Scale); (iii) disability (Schwab and England Activities of Daily Living (ADL) Scale); (iv) comorbid conditions (Charlson co-morbidity index); (v) cognitive function (Mini Mental State Examination (MMSE) test). The demographic information of patients in baseline visit includes: (i) age; (ii) sex; (iii) smoking history in pack-years; (v) whether lives alone or not. 



**Outcome**

The event of interest is institutionalisation, which is defined as entry to a nursing home, residential home, or long-stay hospital ward. Hence, the date of institutionalisation is the date of entry into the institution. Place of residence was ascertained at each annual visit and at death. If the participant had been institutionalised the date was confirmed with the patient, the next of kin, or the institution itself. Patients were censored at the date when they were lost to follow, died without being institutionalised, or if they were alive at the end of follow-up and were not admitted to an institution. The follow-up time is the difference between the event/censoring time and the date of baseline visit. In total, there are  556 participants,  298 patients with PD or atypical parkinsonism, and 123 events in PD or atypical parkinsonism group and 43 in control group.

**Statistical analysis**

*Kaplan-Meier survival curves*


The Kaplan-Meier (KM) estimated survival function was plotted and used to calculate the median time to institutionalization of patients with PD or atypical parkinsonism and those in the control group. KM estimate is a nonparametric estimation [@KM] of cumulative survival probability., It gives the survival probability of every unique event time. At each time the event occurs for a patient, the cumulative survival probability is calculated by one minus the number of individuals who enter an institution divided by the number of patients at risk. Therefore, at time zero, the cumulative survival probability is always one and the survival probability will change over time. Separate KM survival curves were plotted to show survival over time, and incidence rates were calculated for each of these three groups. The incidence rates per 100 person-years were calculated as the total number of events in each group divided by 100 times the sum of follow- up time (year) in each group.

*Multiple imputation*

This study aims to identify risk factors for institutionalization for PD and atypical parkinsonism patients. There are 7 individuals missing the data of smoking history in pack-years (2.2%), 26 individuals missing the data of MMSE test score at the baseline (8.3%) and 58 individuals missing the data of MMSE test score at year one (18.4%). Since the data was incomplete, the cluster of missing data was plotted to see what variables tend to miss together. Then the pattern of missing data was plotted to see how the missing data is connected. There are three types of missing data [@misstype], missing completely at random (MCAR), missing at random (MAR), and missing not at random (MNAR). MCAR is when the probability of being missing doesn't depend on the observed variable. MAR occurs when the probability of being missing depends on observed information. If the reason of missing probability is unknown, then the data is MNAR. In this study, the assumption of missing data was missing at random (MAR), hence we used the multiple imputation to impute patients’ missing data by predictive mean matching method with multivariate imputation by chained equations (MICE) algorithm [@mice]. The multiple imputation procedure uses observed data to impute the missing data and the predictive mean matching method does not require assumptions around the distribution of missing variables. For example, patients without the MMSE test score at baseline can be imputed by their MMSE test score at year one and all other observed information of them. The imputed data draws repeating from the posterior probability distribution of missing data given the observed information. There were 33 patients (10.5%) missing with MMSE test score at baseline and smoking history in pack-years, hence, the imputation times was 10. Because the MICE algorithm uses Markov chain Monte Carlo (MCMC) method, we plotted the parameters against the iteration number to determine whether the MICE algorithm has converged [@convergence]. Convergence was diagnosed as the plot showed no trend. The goodness-of-fit of the multiple imputation model were checked by the density plot of imputed versus observed data.
 

Cox regressions were built to identify the hazard ratio of potential risk factors. Cox regression [@Cox] is a semiparametric model for a time-to-event outcome. The model is composed of the baseline hazard function and relative hazard function, $h(t|X)=h_{0}(t)exp(X\beta)$. The relative hazard function is an exponential formula of risk score which is a summation of the model coefficients multiplied by the value of the patient’s predictor. The Cox model has the assumption of proportional hazards, which means the hazard ratio of predictors does not change over time. But it does not assume the baseline hazard function is constant. The Cox regression can give the hazard ratio of predictors, which is the ratio of hazard rates between two groups and the higher the hazard ratio the higher the risk of the event over time. 


Since the UPDRS part 3 and Hoehn and Yahr Scale are both Parkinson’s rating scales that are used to assess patient symptoms, they may have a similar predictive effect. Therefore, we first built three Cox regression models: (i) Full model; (ii) Model without Hoehn and Yahr Scale; (iii) Model without UPDRS part 3. Then, we compared these three models using the following performance measures: Nagelkerke R-squared, Akaike information criterion (AIC) and Bayesian information criterion (BIC). $R^2$ gives the proportion of the variability of the dependent variable that can be explained by the independent variable. However, $R^2$ is too sensitive in Cox regression as it is binary outcome and involves time variable [@R1;@R2]. Hence,we used Nagelkerke R-squared [@Nagelkerke]. The equation of Nagelkerke R-squared is $R^2_{N}=\frac{R^{2}_{LR}}{R^2_{max}}$ and $R^{2}_{LR}= 1-\lambda^{2/n}$, where $\lambda$ is the null model likelihood divided by the fitted model likelihood and $R^2_{max}$ is the maximum of $R^{2}_{LR}$. It is used to measure the goodness of fit of the model and ranges from zero to one. $R^2_{N}$ of one indicates model is perfectly fit. AIC also measures the goodness of fit of model and the lower AIC value means the model fits better [@AIC1;@AIC2]. BIC [@BIC1] and AIC are both used to compare model with different numbers of parameters. Model with lower BIC value is more likely to be the true model. AIC tends to choose a bigger model, while BIC tends to choose a small model because BIC penalizes model complexity more heavily. Hence, AIC and BIC are used together to choose model [@BIC2]. The formula is $-2logL+kp$, where L is the likelihood function and p is the number of parameters in the model, and k is 2 for AIC and log(n) for BIC.


After selecting the model, the model assumption of proportional hazards was checked by Schoenfeld residuals method. The Schoenfeld residual [@Schoenfeld1] is the observed value of the covariate for patient minus the expected value of this patient. And under the proportional hazards assumption, plot of the scale Schoenfeld residuals of each predictors over time should not have trends. If the Schoenfeld residual plot can fit a straight line, then the p-value of the hypothesis test of proportional hazards assumption should be larger than 0.05 [@Schoenfeld2]. Then, we accessed the functional form of each continuous predictor by plotting Martingale residual against continuous predictor [@Martingale1]. If the fitted model is correct, these residuals should symmetrically distributed with expected value 0 [@Martingale2]. However, the plot showed trends in “age” and “Schwab & England ADL score”. To determine whether they need to be transformed or not, we compared models with and without restricted cubic splines (RCS) [@RCS] by $R^2$, AIC and BIC. Finally, we presented the hazard ratios and confidence interval of estimate effects of predictors in model by figure and table. 

 
The statistical analysis is performed by R verison 4.0.5 and R packages [@xtable;@dplyr;@tidyr;@Hmisc;@gridExtra;@readstata13;@survival;@survminer;@rms;@ggplot2]. 



## Results



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,eval = T,warning = F,message = F)
```

```{r library}
library(xtable)
library(dplyr)
library(tidyr)
library(knitr)
library(gridExtra)
library(Hmisc)
library(readstata13)
library(survival)
library(survminer)
library(rms)
library(skimr)
library(finalfit)
library(mice)
library(ggplot2)
library(GGally)
library(png)
library(bookdown)
library(foreign)
library(kableExtra)
library(patchwork)#arrange survival plot
library(sjlabelled)

#install.packages('patchwork')
```

```{r options}
options( tinytex.verbose = TRUE)

```

```{r }
data<-read.dta13("updata3.dta",nonint.factors = T,missing.type = T)
data[is.na(data)] <- NA  #575
```

```{r data}

attach(data)

data<-data %>%
  mutate(t=dateinstitution)%>%           
  mutate(t=coalesce(t,datelost))%>%
  mutate(t=coalesce(t,datedeath))%>%
  mutate(t=coalesce(t,datelastknownalive))

# sum(is.na(data$t))

data$cens<-ifelse(is.na(dateinstitution),0,1)


data$tt<-as.Date(as.character(data$t), format="%Y-%m-%d")-
                as.Date(as.character(data$datevisitbl), format="%Y-%m-%d")

data$tt<-as.numeric(data$tt)

data$year<-data$tt/365.25


# str(data)

data$sex<-factor(data$sex,
                levels = c(0,1),
                labels = c("male", "female"))

data$livesalone<-factor(data$livesalone,
                        levels = c(0,1),
                        labels = c("no", "yes"))


data<-data%>%
  filter(data$tt>0) #556

```




```{r data2}

data2<- subset(data,!(diagnosiscat %in% c("Control")))

data2$group<-ifelse(data2$diagnosiscat=="Atypical parkinsonism",1,0)   # (,yes,no) Atypical parkinsonism=1, Parkinson's disease=0


data2$group<-factor(data2$group,
                    levels = c(0,1),
                    labels = c("Parkinson's disease","Atypical parkinsonism"))

```


**Describe of data**

```{r datasum}

data.sum<-read.csv("demographic summary.csv")%>%
  rename("Characteristic"=ï..Characteristic,"Parkinson’s disease (N=201)"=Parkinsonâ..s.disease..N.200.,"Atypical parkinsonism (N=114)"=Atypical.parkinsonism..N.98.,"Control (N=260)"=Control..N.258.)

data.sum%>%
  kable(caption='\\label{tab:datasum} Summary demographic information of participants.',booktabs=T,linesep="") 






```

Table \@ref(tab:datasum) shows that 70 (35%) PD patients, 53 (54.1%) atypical patients and 43 (16.7%) patients in control entered institutionalisation. And 401 deaths occurred (140 in control group) and 6 were lost to follow up (3 in PD and 3 in control group). The population in this study is elderly participants as most of them are over 70 years old and the sex ratio between male and female are around 1.5. Both age and sex distributions are similar in these three groups. There are more individuals have smoking habits and living alone in the control group, while heavier smoker in atypical parkinsonism group. And less living alone patients in PD group than others. In short, the demographic information of participants are very similar among these three group with slightly difference in smoking habit and whether living alone. 


```{r eval=FALSE}

# MMSEKM,fig.align = "center",fig.cap = "\\label { } Kaplan-Maier survival curves of MMSE groups.", fig.pos ='H'

data$mmsecat<-cut(data$mmsebltotal,c(0,23,27,30),labels = c("<24","24-27","28-30"))

MMSEKM<-survfit(Surv(tt/365.25,cens)~mmsecat,data=data)

ggsurvplot(MMSEKM,ylab="Survival probability of institutionalisation",xlab="Survival years",data,legend="top",surv.median.line = "hv",legend.labs=c("MMSE:<24","MMSE:24-27","MMSE:28-30"))


```

The fig \@ref(fig:MMSEKM) shows that the KM curves of three MMSE groups are all separated which means it is reasonable to divided MMSE score in three different groups. Patients with probable dementia (MMSE score less than 24) has the shortest follow-up time to the institutionalisation. 




 
**Kaplan-Meier estimation** 


```{r reKM}

data$cens.followup<-1-data$cens

reverse.KM<-survfit(Surv(tt/365.25,cens.followup)~diagnosiscat,data=data)

# KM2

data.frame(groups= c("Parkinson's disease","Atypical parkinsonism","Control"),n=c(200,98,258),median=c(9.26,4.41,10.78),LCL=c(8.30,3.33,10.40),UCL=c(11.08,5.78,11.07))%>%
  kable(caption='Reverse Kaplan-Meier estimate of the median follow-up time to institutionalisation.',booktabs=T,linesep="",digits=2) %>%
  footnote(alphabet = c ("The time is year",
               "LCL is the 95% lower confidence limits;", 
               "UCL is the 95% upper confidence limits;"
               ))


```
Table \@ref(tab:reKM) shows the median follow-up time for PD patients (9.26 years) is 2.1 times longer than atypical parkinsonism patients (4.41 years). 


```{r KMplot, fig.align = "center",fig.cap = "\\label { } Kaplan-Maier survival curve.", fig.pos ='H'}

KM<-survfit(Surv(tt/365.25,cens)~diagnosiscat,data=data)


ggsurv<-ggsurvplot(KM,
           ylab="Survival probability of institutionalisation",
           xlab="Survival years",
           data,legend="top",
           surv.median.line = "hv",
           risk.table = T,
           risk.table.y.text.col = F,
           risk.table.y.text = T,
           risk.table.col = "strata",
           linetype = "strata",
           ggtheme = theme_bw(),
           legend.labs=c("Parkinson's disease","Atypical parkinsonism","Control"))

print(ggsurv, 
      surv.plot.height = 0.7,
      risk.table.height = 0.3)
```
The fig \@ref(fig:KMplot) shows that at the beginning the survival probability of entering the institutionalisation is overlap between PD group and control group, however the survival probability for atypical parkinsonism group drops very quickly. The KM curve of PD patients is above atypical parkinsonism KM curve all the time. And number at risk table shows only 2 patients didn’t enter the institutionalisation in atypical parkinsonism group.  Hence, patients with PD had longer independent living time than patients with atypical parkinsonism. And the median survival time for PD patients entering long-term care facility (11.8 years) was 3.4 times longer than atypical parkinsonism patients (3.5 years).


**The incidence rates**

```{r incidence}



data%>%
  group_by(diagnosiscat)%>%
  summarise('Incidence rate'=paste(round(sum(cens)/as.numeric(sum(year))*100,1),' per 100 person-years'))%>%
  rename(Group=diagnosiscat)%>%
  kable(caption=' The incidence rate of institutionalisation.',booktabs=T,linesep="") 



```

Table \@ref(tab:incidence) shows that atypical parkinsonism group (27.6 per 100 person-years) has 5.3 times higher incidence rate of institutionalisation than PD group (5.2 per 100 person-years). Therefore, atypical parkinsonism patients have higher risk of entering institutionalization than PD patients.


**Multiple imputation**


```{r missplot, fig.align = "center",fig.cap = "\\label { } Cluster analysis showing which predictors tend to be missing on the same patients.", fig.pos ='H'}

data3<-data2

data3<-rename(data3,c("Age at baseline visit"=agebl, "Age at diagnosis of parkinsonian syndrome"=agedx,"Age at onset of of parkinsonian symptoms"=agesymptomonset,"Smoking history in pack-years"=packyears,"ever smoke"=eversmok,"Lives alone"=livesalone,"Charlson co-morbidity index"=charlsonbl,"UPDRS part3 (motor) score"=updrsblpart3,"Hoehn and Yahr Scale"=hybl,"Schwab and England ADL Scale"=sebl,"MMSE at baseline"=mmsebltotal,"MMSE at year 1"=mmseyr1total))



plot(naclus(data3[,c(6:18)]))

```

From the fig \@ref(fig:missplot), we can see that Mini Mental State Examination test score at baseline and Mini Mental State Examination test score at year 1 are likely to be found missing in the same patients. 


```{r misspattern, fig.align = "center",fig.cap = "\\label { } Missing data pattern.", fig.pos ='H'}

explanatory<-c("Smoking history in pack-years","MMSE at baseline","MMSE at year 1")
dependent<- c("cens","tt")

mispattern<-data3 %>% 
  missing_pattern(explanatory)

```
The fig \@ref(fig:misspattern) is the visualisation of this missing pattern output, where the blue cubic means observed and red cubic means missing. From the fig \ref{fig:misspattern}, we can see that 24 patients missing with the minimental state examination score at baseline and 6 patients missing with smoking history in pack-years. The missing pattern is general and connected through the first row. The first column shows the frequency of each pattern and the number of missing entries in each pattern shows in the last column. For example, the first row means there are 237 patients in the dataset with all three observed variables. 




```{r pmm}

pmm<-mice(data2[,c(6,9:18,25:26,28)],method="pmm",seed=81420,m=10,print=F)


```


```{r convergence, fig.align = "center",fig.cap = "\\label{ } Mean and standard deviation of the synthetic values plotted against iteration number for the imputed data.", fig.pos ='H'}
# pmm$visitSequence

plot(pmm)




```
From fig \@ref(fig:convergence), we can see that there are no obvious trends. Hence, the MICE algorithm has converged.


```{r densityplot1,fig.height=5, fig.width=10, fig.align = "center",fig.cap = "\\label{ } Density plot of the observed and imputed data of MMSE test score at baseline ", fig.pos ='H' }

densityplot(pmm,~ mmsebltotal,xlab=" Mini Mental State Examination test score at baseline (red line is imputed data and blue line is observed data)" )



```

From the fig \@ref(fig:densityplot1), we can see that the distribution of observed and imputed data is similar.


```{r mmsecat}

long<-complete(pmm, 'long', include=TRUE) #include=T means, the long contains original data and 10 imputed data.
                                          #if include=FALSE or not specified, Error in as.mids(long) : Original data not found.


N <- nrow(long)
for (i in 1:N){
  if(!is.na(long$mmsebltotal[i])){
    if (long$mmsebltotal[i] <= 23 ){
      long$mmsecat[i] = 0
    }else if (long$mmsebltotal[i] <= 27 && long$mmsebltotal[i] > 23){
      long$mmsecat[i] = 1
    }else if (long$mmsebltotal[i] <= 30 && long$mmsebltotal[i] > 27){
      long$mmsecat[i] = 2
    }
  }
}




long$mmsecat<-factor(long$mmsecat,
                        levels = c(0,1,2),
                        labels = c("<24", "24-27", "28-30"))



pmm.new<-as.mids(long)



```



```{r age10}



long$age10<-long$agebl/10
                        
long$packyears10<-long$packyears/10

long$updr3.10<-long$updrsblpart3/10

long$sebl10<-long$sebl/10

pmm.new<-as.mids(long)



```


**Identify the risk factors**



```{r cox models,results = 'hide'}

dd<-datadist(long)
options(datadist='dd')

modela<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+hybl+sebl+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new)  #full

modelb<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+sebl+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) #without hybl

modelc<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) #without updrsblpart3

```

```{r ,eval=FALSE}
modela
modelb
modelc

```



```{r modelselect}

AIC1<-AIC(modela)  #full
AIC2<-AIC(modelb)  #without hybl
AIC3<-AIC(modelc)  #without updrs


BIC1<-BIC(modela)  #full
BIC2<-BIC(modelb)  #without hybl
BIC3<-BIC(modelc) #without updrs


data.frame(Model=c("Full Model","Model without Hoehn & Yahr disease stage","Model without UPDRS part 3 (motor) score") , 'Nagelkerke R-squared' = c(0.36,0.36,0.36),AIC = c(AIC1,AIC2,AIC3), BIC = c(BIC1,BIC2,BIC3))%>%
  kable(caption='Model tests and discrimination indexes.',booktabs=T,linesep="",digits=2)


```
Table \@ref(tab:modelselect) shows that the $R^{2}_{N}$ of these three models are the same and the model without Hoehn & Yahr disease stage has the lowest AIC and BIC value. Due to only has 123 events in PD or atypical parkinsonism group and to avoid over fitting, we choose the model without Hoehn & Yahr disease stage. 





```{r label}

labs<-c("Beta(t) for Age",
        "Beta(t) for Sex","Beta(t) for Smoking history in pack-years",
        "Beta(t) for Lives alone",
        "Beta(t) for Charlson co-morbidity index","Beta(t) for UPDRS part 3 (motor) score",
        "Beta(t) for Schwab and England ADL Scale",
        "Beta(t) for Mini Mental State Examination test",
        "Beta(t) for Diagnosis")

data.label<-set_label(data2[,c(6,9,10,12:14,16,17,28)],label=labs)


colnames(data2)

```

```{r testcox}


pmm.new_names <- paste("pmm.new",1:10,sep="")  # create names first

for (i in 1:10){
    pmm.new_df<-complete(pmm.new,i)
    assign(pmm.new_names[i], pmm.new_df)
}


test_names<- paste("test.ph",1:10,sep="")   # create names first

for (i in 1:10){
    test.ph_df<-cox.zph(coxph(Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+sebl+mmsecat+group,data=get(paste0("pmm.new",i))),transform = "km")   # data=get(paste0("pmm.new",i))
    assign(test_names[i], test.ph_df)
}


test.ph<-unlist(test.ph10[["table"]])

test.ph<-as.data.frame(test.ph)
  
rownames(test.ph)<-c("Age","Sex","Smoking history in pack-years","Lives alone","Charlson co-morbidity index","UPDRS part 3 (motor) score","Schwab and England Activities of Daily Living (ADL) Scale","Mini Mental State Examination (MMSE) test","Atypical parkinsonism vs PD group","Global") 


test.ph %>%
   kable(caption='\\label{tab:testcox} PH assumption test results.',booktabs=T,linesep="",digits=2)


test.ph.plot<-cox.zph(coxph(Surv(tt/365.25,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+sebl+mmsecat+group,data=pmm.new10),transform = "km") 




plot(test.ph.plot,ylab = label(data.label))




p1<-ggcoxzph(test.ph.plot,var=1,ylab ="Beta(t) for Age",caption = NULL)+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p2<-ggcoxzph(test.ph.plot,var=2,ylab = "Beta(t) for Sex",caption = NULL)+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p3<-ggcoxzph(test.ph.plot,var=3,ylab = "Beta(t) for Smoking history in pack-years",caption = NULL )+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p4<-ggcoxzph(test.ph.plot,var=4,ylab = "Beta(t) for Lives alone",caption = NULL )+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p5<-ggcoxzph(test.ph.plot,var=5,ylab = "Beta(t) for Charlson co-morbidity index",caption = NULL)+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p6<-ggcoxzph(test.ph.plot,var=6,ylab = "Beta(t) for UPDRS part 3 (motor) score",caption = NULL )+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p7<-ggcoxzph(test.ph.plot,var=7,ylab = "Beta(t) for Schwab and England ADL Scale",caption = NULL )+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p8<-ggcoxzph(test.ph.plot,var=8,ylab = "Beta(t) for Mini Mental State Examination test",caption = NULL )+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p9<-ggcoxzph(test.ph.plot,var=9,ylab = "Beta(t) for Diagnosis",caption = NULL)+
  theme_survminer(font.main = 8,font.y = 8,font.x = 8,axis.text.x = element_text(size = 1))

p1<-p1[["1"]]
p2<-p2[["2"]]
p3<-p3[["3"]]
p4<-p4[["4"]]
p5<-p5[["5"]]
p6<-p6[["6"]]
p7<-p7[["7"]]
p8<-p8[["8"]]
p9<-p9[["9"]]

p1+p2+p3
p4+p5+p6
p7+p8+p9



```



Table \@ref(tab:testcox) shows the results of proportional hazards (PH) assumption testing, we can see that all variables’ p-value is larger than 0.05. Hence there is not enough evidence to reject the null hypothesis that the proportional hazards are constant and the PH assumption is valid.

```{r martingale,fig.align = "center",fig.cap = "\\label{ } Martingale residuals versus continuous predictors.", fig.pos ='H'}

pmm.new1<-complete(pmm.new,1)
rr.0<-residuals(coxph(Surv(tt,cens)~1,data = data2),type="martingale")

pmm.new1$rr.0<-rr.0


plot1<-ggplot(pmm.new1,aes(agebl, rr.0)) + 
         geom_point(cex=1) +
         geom_smooth(method = "loess", se = T)+
         xlab("Age")+
         ylab("residuals")+
         ylim(c(-2,2))



plot2<- ggplot(pmm.new1,aes(packyears, rr.0)) + 
         geom_point(cex=1) +
         geom_smooth(method = "loess", se = T) +
         xlab("Smoking history in pack-years")+
         ylab("residuals")+
         ylim(c(-2,2))   

plot3<-ggplot(pmm.new1,aes(updrsblpart3, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        xlab("UPDRS part 3 (motor) score")+
        ylab("Martingale residuals")+
        ylim(c(-2,2))


plot4<- ggplot(pmm.new1,aes(sebl, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        xlab("Schwab and England ADL Scale")+
        ylab("residuals")+
        ylim(c(-2,2))

plot5<- ggplot(pmm.new1,aes(charlsonbl, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        xlab("Charlson co-morbidity index")+
        ylab("residuals")+
        ylim(c(-2,2))


grid.arrange(plot1,plot2,plot3,plot4,plot5, ncol=2, nrow=3)





```
From the \@ref(fig:martingale) we can see that "Smoking history in pack-years" and "Charlson co-morbidity index" are both seems to be symmetrically distributed around $\pm 1$ and with flat blue fitted line around zero. And the scatter points of "Hoehn & Yahr Scale" seems symmetrically distributed around $\pm 0.5$ with expected value of zero. For the scatter plot of "Schwab & England ADL Scale", there appears to be a slightly decreasing trend  while for "Age" the line is slightly going up. 


```{r rcs,results = 'hide'}

modelc1<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+updrsblpart3+sebl+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age as 3 knot rcs


modelc2<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+rcs(updrsblpart3,3)+sebl+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with updrsblpart3 as 3 knot rcs


modelc3<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+rcs(sebl,3)+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with sebl as 3 knot rcs Choose


modelc4<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+rcs(updrsblpart3,3)+sebl+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age and updrsblpart3  as 3 knot rcs


modelc5<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+updrsblpart3+rcs(sebl,3)+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age and sebl as 3 knot rcs

modelc6<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+rcs(updrsblpart3,3)+rcs(sebl,3)+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with and updrsblpart3 and sebl as 3 knot rcs


modelc7<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+rcs(updrsblpart3,3)+rcs(sebl,3)+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age and updrsblpart3 and sebl as 3 knot rcs





```


```{r ,eval=FALSE}
modelc1
modelc2
modelc3
modelc4
modelc5
modelc6
modelc7

```


```{r modelselect2}
AIC.b<-AIC(modelb)    # linear


BIC.b<-BIC(modelb)   # linear


AICb1<-AIC(modelc1)
AICb2<-AIC(modelc2)
AICb3<-AIC(modelc3)
AICb4<-AIC(modelc4)
AICb5<-AIC(modelc5)
AICb6<-AIC(modelc6)
AICb7<-AIC(modelc7)

BICb1<-BIC(modelc1)
BICb2<-BIC(modelc2)
BICb3<-BIC(modelc3)
BICb4<-BIC(modelc4)
BICb5<-BIC(modelc5)
BICb6<-BIC(modelc6)
BICb7<-BIC(modelc7)


# restricted cubic splines (RCS)

data.frame(Model=c("Linear model","RCS model with 3 knots in age","RCS model with 3 knots in UPDRS part 3 (motor) score","RCS model with 3 knots in  Schwab and England ADL Scale","RCS model with 3 knots in age and UPDRS part 3 (motor) score","RCS model with 3 knots in age and Schwab and England ADL Scale","RCS model with 3 knots in UPDRS part 3 (motor) score and Schwab and England ADL Scale","RCS model with 3 knots in age and UPDRS part 3 (motor) score and Schwab and England ADL Scale") , 'Nagelkerke R-squared' = c(0.36,0.36,0.36,0.36,0.36,0.36,0.36,0.36),AIC = c(AIC.b,AICb1,AICb2,AICb3,AICb4,AICb5,AICb6,AICb7), BIC = c(BIC.b,BICb1,BICb2,BICb3,BICb4,BICb5,BICb6,BICb7))%>%
  kable(caption=' Compare model with and without Restricted cubic splines (RCS).',booktabs=T,linesep="",digits=2)

```
Table \@ref(tab:modelselect2) shows that there is only little difference between $R^2$, AIC and BIC value of models with or without RCS terms. And the linear model has the lowest BIC value and the second lowest AIC value. Therefore, since the simplest model is just as good as the models with RCS terms, we choose the linear model.




```{r coxout}

sumrcs2<-data.frame(summary(modelb))
sumrcs2.new<-sumrcs2[grep("Hazard.Ratio",rownames(sumrcs2)),]


row.names(sumrcs2.new)<-c("Age","Smoking history in pack-years","Charlson co-morbidity index","UPDRS part 3 (motor) score","Schwab and England Activities of Daily Living (ADL) Scale","Sex - female:male","livesalone - yes:no","Mini Mental State Examination (MMSE) test - <24:28-30 ","Mini Mental State Examination (MMSE) test - 24-27:28-30","Group - Atypical parkinsonism:PD")


#sumrcs2.new%>%
#  select(Effect,Lower.0.95,Upper.0.95)%>%
#  rename("Hazard ratio"=Effect,"The 95% lower confidence limits"=Lower.0.95,"The 95% upper confidence limits"=Upper.0.95)%>%
#  kable(caption='Summary of estimated effects.',booktabs=T,linesep="",digits=2)


read.csv("Hazard ratio.csv")%>%
  rename("Characteristic"="ï..Characteristic","Hazard ratio and the 95% confidence interval"="Hazard.ratio.with.95..confidence.interval")%>%
  kable(caption='Summary of estimated effects.',booktabs=T,linesep="",digits=2)






modelb.new<-fit.mult.impute((Surv(tt,cens)~age10+sex+packyears10+livesalone+charlsonbl+updr3.10+sebl10+mmsecat+group),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new)


modelb.new


  
```

```{r }

round(exp(0.2024),2)

round(exp(-0.0163-1.96*0.0692),2)

round(exp(-0.0163+1.96*0.0692),2)


```


Table \@ref(tab:coxout) shows that the significate risk factors were increasing age for 10 yeas [HR=1.89 (1.43,2.51)], MMSE score less than 24 [HR versus MMSE>27 =2.60 (1.43, 4.72)], and being diagnosed as atypical parkinsonism [HR versus PD = 3.02 (1.84,4.96)]. Since their 95% confidence interval didn’t include one, they were independently associated with institutionalisation. 

Patients with less than 24 score in MMSE indicates poor cognition and had 2.6 times higher risk of entering the institutionalisation than patients with normal cognition. And the chance for atypical parkinsonism patients being admitted into institutionalisation was 3 times higher than PD patients. Patients’ risk of being entered into the institutionalisation increase with age, however, patient’s sex and whether lives alone were both not associated with this risk as their 95% hazard ratio confidence interval both include one. 

For the patients’ clinical examinations of parkinsonian impairment (UPDRS part3 (motor) score), disability (Schwab and England Activities of Daily Living (ADL) Scale), and comorbid conditions (Charlson co-morbidity index) were all not statistically significant as their 95% hazard ratio confidence interval all include one. The point estimate hazard ratio indicated that less independence patients (lower percentages in Schwab and England ADL Scale) and more serious disability patients had higher chance of being admitted into institutionalisation. However, patients with more serious comorbid conditions had less chance being entered into long term care facilities.


```{r coxoutplot,out.width="70%",fig.align = "center",fig.cap = "\\label{ } Hazard ratios and confidence bars for risk factors in model.", fig.pos ='H'}


text<-"Notes:\nagebl is age; packyears is smoking history in pack-years;\ncharlsonbl is Charlson co-morbidity index; hybl is Hoehn and Yahr Scale;\nsebl is Schwab and England Activities of Daily Living (ADL) Scale;\nmmsecat is Mini Mental State Examination (MMSE) test."
par(mar = c(8,5,3,3))
plot(summary (modelb),main= '',log = T)
mtext(text, side = 1, line = 5, cex = 1.1, adj = 0) 


data%>%
  group_by(diagnosiscat)%>%
  filter(!is.na(datelost))%>%
  count()

163/200

```

The \@ref(fig:coxoutplot) shows the hazard ratio in continuous characteristic. We can see that every 10 years older for patients, double the risk of entering the nursing home. However, the risk of institutionalisation between patients who smoked 22 packs in a year and patients without smoking habit was no significant difference. 



## Discussion 


This study found that patients with atypical parkinsonism have 5.3 times higher incidence rate (27.6 per 100 person-years) of entering the long-term care facilities and 4.7 times shorter median follow-up time to institutionalisation (2.5 years) than patients with Parkinson’s disease (PD). The highest hazard ratio risk factor was probable dementia (HR=3.1), followed with 10 years ageing (HR=2.0) and probable cognitive impairment (HR=1.2). And these three risk factors were three statistical significance as their 95% confidence interval didn't include one. Female patients who live alone with heavy smoking habit were more likely need long-term care. And patients with more serious disability or disease stage had higher chance of admitting into nursing home. However, patients with more serious comorbid conditions tended to have less chance of entering institutionalisation. 


The PINE study is population-representative cohort of newly diagnosed patients with degenerative or vascular parkinsonism in Aberdeen, Scotland. Since study follows the patients from diagnosis to enter the institutionalisation, there is less bias between the study population. And PINE study is a long-term follow up study (the median follow-up time in PD group is 9.26 years, atypical parkinsonism group is 4.41 years, and control group is 10.78 years) with 575 patients and only 6 patients lost follow up. Compared to the previous four-year prospective study of 178 community-dwelling subjects in PD [@pred2000], our study has longer follow-up time and 3.23 times more patients involved. However, due to the degenerative or vascular parkinsonism is more common in elderly than young people. The PINE study is a elderly cohort study, the median age is 76.8 years old and only 14 patients' age is under 50 years old. Therefore, this study results may not represent the institutionalisation risk of young patients. 


This study used multiple imputation method to imputed  patients’ missing data: 7 missing data in smoking history (pack-years), 26 missing data in MMSE test score at the baseline and 58 missing data in MMSE test score at year one. Compared to single imputation, multiple imputation is more robust. In the previous studies [@pred2000; @pred2015], they didn't report the missing data and what kind of methods they used to deal with missing data. And we used cox regression to investigate the risk factor of institutionalisation. Because the PD is a slowly progressive disorder and from diagnosis to entering nursing home usually take years. Therefore, during the follow-up study, it may lost to follow-up patients or patients may die before entering long-term care facilities. And cox regression is a model from time-to-event outcome. Compared to logistic regression, the cox regression includes the time variable and can represent event/censor data. Therefore, for the study that has more chance of censoring, the cox regression has more advantage than logistic regression. In the nationwide retrospective cohort study [@pred2015], they used logistic regression to analyse a large cohort (469005 patients) and didn't report how many missing or lost to follow-up. And in this study, we tried to keep numeric data in continuous instead setting cut-off points. Because setting cut-off points (no based-on clinician advices) in continuous data will lost information and the model will have more bias. For example, the prospective study in Norway [@pred2000] set cut-off points in age and found that PD patients who older than 72 years old will have 3.5 times higher risk of nursing home admission than those younger or equal to 72 years old. Their model result means for patient who is just turned 73 and who was in last month of 72, their risk is can 3.5 times different. While in our study, we kept age as continuous data and found every 10 years ageing the risk of entering nursing home will be double. Keeping data as continuous can allow clinician to make more precisely predict and the prediction model is less bias. Furthermore, in our study, we chose the variable based on clinician suggestions instead of selecting statistics significant variables by statistics methods. In the Norway study [@pred2000], they used bivariate Cox analyses to select variables and transformed continuous data into dichotomous data, hence their 95% confidence interval of relative risk (in the model result) was very wide. Our model is more robust due to the 95% confidence interval of hazard ratio is narrow. The only continuous data we changed into category is MMSE. The cut-off points in MMSE are based on the clinician suggestion and and the mean MMSE score in normal healthy adults showed in previous study [@MMSE2]. In order to make model more robust, we kept the MMSE as continuous data when we performed multiple imputation and only changed it as category when built Cox regression model. 


In our study, we found that the risk of female entering nursing home is only slightly higher than male (HR=1.12) and sex is not a significant risk factor (95% CI include one) to the institutionalisation. This result also found in the previous study in Norway [@pred2000]. However, the incidence rate of PD for male is higher than female [@IncidencePD] and male-to-female ratio is around 3:2 [@Lancet2015]. Therefore, male has higher risk of PD, but same risk as female to the nursing home. And many studies showed that patients with smoking habit have less risk of PD than those who never smoke [@EpidemiologyPD]. But the biological basis of this association is poorly understood, one possible reason is nicotine has a neuroprotective action and may stimulate dopamine release [@SmokePD]. But, we found only heavy smoker will have slightly significant higher risk of entering the long-term care facilities because the hazard ratio between non-smokers and patients who smoke 23 pack-years is 1.18. It means if heavey smoker patients do not reduce their smoking times after diagnosed with PD or atypical parkinsonism, they will have higher risk of entering nursing home. Different to previous studies found that lives alone is a risk factor to institutionalisation [@pred2000], we found living alone is not a statistical significant risk factor as the 95% confidence interval include one. And in PINE study, the number of patients who live with others (97 individuals) is 2.3 times more than those who live alone (43 individuals) in institutionalisation. Hence, for PINE study, living alone patients are less common in institutionalisation. This may because the local social services system provide help and extend patients' living at home time. Therefore, patients only enter the nursing home when the social care cannot take care of them. Furthermore, the measurement of Parkinson's symptoms progression (Hoehn and Yahr Scale) and independence of daily activities (Schwab and England ADL Scale) are both not significant while the cognitive function (MMSE test) is the most important risk factor shows that despite patients may have disability in life but as long as patients' cognitive is still normal, the chance of they can take care by others or by themselves is higher than those who have probable dementia. Therefore, treatment of cognitive impairment may extend patients' survival time to long-term care facilities. In short, our study found that cognitive impairment and old age are the most important risk factor for patients entering the institutionalisation. And for further study, it can investigate the relationship between local social services for living alone patients and the risk of entering in long-term care facilities. For clinicians, if a newly diagnosed atypical parkinsonism patient has cognitive impairment, then this patient has highest risk of entering the institutionalisation and should plan management earlier. 








 















```{r eval=FALSE}
my_skim <- skim_with(base = sfl(n = length))

data2%>%
  group_by(diagnosiscat)%>%
  dplyr::select(agebl)%>%
  my_skim()
  


data$mmsecat<-cut(data$mmsebltotal,c(0,23,27,30),labels = c("<24","24-27","28-30"))

MMSEKM<-survfit(Surv(tt/365.25,cens)~mmsecat,data=data)

ggsurvplot(MMSEKM,ylab="Survival probability of institutionalisation",xlab="Survival years",data,legend="top",surv.median.line = "hv",legend.labs=c("MMSE:<24","MMSE:24-27","MMSE:28-30"))



aloneKM<-survfit(Surv(tt/365.25,cens)~livesalone,data=data)

ggsurvplot(aloneKM,ylab="Survival probability of institutionalisation",xlab="Survival years",data,legend="top",surv.median.line = "hv")

data2%>%
  filter(data2$cens==1)%>%  #cens=1 event happened
  group_by(livesalone)%>%
  count(livesalone)








```





# Reference


