---
title: "Dynmatic predtion model in parkinson disease"
author: "Yan Li"
header-includes:
- \usepackage{floatrow}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \floatsetup[table]{capposition=top}
- \floatplacement{table}{H}
output:
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
        scroll_highlight: yes
      before: null
      after: null
    toolbar:
      position: fixed
    edit : null
    download: [["book.pdf", "Download PDF version"]]
    search: yes
    fontsettings:
      theme: white
      family: sans
      size: 2
    sharing:
      facebook: yes
      github: no
      twitter: yes
      linkedin: no
      weibo: no
      instapaper: no
      vk: no
      whatsapp: no
      all: ['facebook', 'twitter', 'linkedin', 'weibo', 'instapaper']
    info: yes
  bookdown::html_book:
    css: toc.css
  bookdown::pdf_book:
    keep_tex: yes
documentclass: book
bibliography: my5.bib
fontsize: 14pt
citation_package: none
csl: harvard-cite-them-right2.csl
nocite: '@*'
---
# Systematic review of prognostic models in Parkinson’s disease

# PD cox regression

## Abstract

When patients with Parkinson’s disease lose their ability to take care of themselves, and carers are no longer able to look after them at home, they will need long-term care. The risk factors for being admitted to a nursing home are unclear and there are few published studies. This work will identify the incidence of institutionalization and  will identify the risk factors for institutionalization using data from a prospective incident cohort of parkinsonism in North East Scotland.


## Introduction

Parkinson’s disease (PD) and atypical parkinsonism are neurodegenerative disorders. As the disease progresses, the need for care in PD patients will increase and many patients will need to enter a long-term care institution, such as a nursing home. No previous studies have reported the incidence of institutionalisation in PD and atypical parkinsonism in a population-representative cohort. Several studies have investigated risk factors for institutionalisation. A previous study suggested that older patients who live alone with motor disability tend to have a higher probability of entering a nursing home [@pred2000]. @pred2015 pointed out that the probability of being admitted into a nursing home was associated with race and sex. In a previous study, researchers found that dementia, hallucinations, cognitive decline, and hip fracture were risk factors of institutionalization for PD patients [@pred1993;@pred2000;@pred2015]. The design of previous studies are population-based prospective study and nationwide retrospective cohort study [@pred2000;@pred2015]. No previous studies have investigated risk factors in a population-representative cohort and none have examined the influence of socioeconomic status on institutionalisation in parkinsonism.

This study will first identify the incidence of institutionalization in PD and atypical parkinsonism. We will then identify the risk factors for institutionalization in PD and atypical parkinsonism. The results from this study will enable clinicians to manage their patients better and earlier and provide more appropriate services to help PD patients.


## Databases

The Parkinsonism incidence in North-East Scotland (PINE) study is a prospective follow-up study that collects annual patient data on survival, disability, and institutionalization [@Field2016]. The pilot study of PINE is from 2002 to 2004 and the main study is from 2006 to 2009. It recruited all patients with a newly diagnosed presumed degenerative or vascular parkinsonian syndrome over 4.5 years from a baseline population of about 315000 patients registered with 37 primary care practices in and around Aberdeen, Scotland. The screening in multiple overlapping searches is limited to the population over 65 years old. Parkinsonism was defined as two or more cardinal motor signs and patients with drug-induced parkinsonism were excluded. There is ongoing yearly follow-up of eligible patients and their carers with linkage to the national death register. The standardized baseline visit at diagnosis and annually thereafter was the same for patients and controls. At each annual review the parkinsonian syndrome was classified by a single consultant neurologist with expertise in movement disorders. The patient diagnosis was assigned by the best fitted criteria. If a patient died, then the diagnosis was made after reviewing all the clinical and imaging information held in their research files and the annual videotaped examinations or from pathology in those who had given consent for post-mortems. For patients who were lost to follow-up, identifying an age-sex matched control from the same primary care practice or a register of elderly people who had taken part in a previous community-based screening project. At each annual visit (and when notification of death was received) the patient’s address was documented. If this was a nursing home or other form of institutionalisation, the date of admission was obtained from the patient, or from the nursing home if the patient was unable to remember. And if the patient died, the details about the date, place and cause of death were extracted from death certificates and primary and secondary care records.

## Study design

The study design is an observational study. 

## Baseline characteristics

In this study, there are four predictor variables to describe patients' basic information and five predictor variables to describe patients' medical information. Patients' basic information includes their age, sex, smoking history in pack-years, whether lives alone or lives with someone. Patients' medical recording includes three rating scales of Parkinson's, one index of patients' death risk, and one test result of patients' mental abilities. 


Parkinson’s rating scales are used to assess patients' symptoms of the condition and evaluate treatment, and clinicians assign a value to it depends on patients' particular feature information. Motor scales and non-motor scales are both important. The Unified Parkinson’s Disease Rating Scale (UPDRS) has four parts, the first three sections evaluate the degree of disability and the fourth section evaluates any complications of treatment. In this study, we used the value from UPDRS part 3 (motor examination). It is scored on a 0-4 rating scale and higher scores means more severe. And the UPDRS is often used with the Hoehn and Yahr Scale and the Schwab and England Activities of Daily Living (ADL) Scale [@Ratingscales]. The Hoehn and Yahr Scale is used to measure the progress of Parkinson’s symptoms and the level of disability. It includes stages 0 to 5, where stage 0 means no signs of disease and stage 5 means need a wheelchair or bedridden. The Schwab and England ADL Scale is to measure patients' independence of daily activities through a percentage figure, the high percentages indicate independence while low percentages indicate dependence. Patients' death risk and mental abilities were also measured. Charlson co-morbidity index predicts one-year mortality risk of hospitalized patients with specific co-morbid conditionsy. It includes 16 conditions and each condition is assigned a weight from 1 to 6 [@Charlson]. The Mini Mental State Examination (MMSE) test is used to diagnose dementia and assess its progression and give indication of severity. The maximum score of MMSE is 30 points, less than 24 points indicates mild dementia. And the MMSE score of Alzheimer's patients drops around two to four points averagely each year [@MMSE]. Hence, if patient's medical history record without the baseline MMSE score, we can use MMSE score on year one to infer the MMSE score on baseline. 


In short, there are nine predictor variables in the Cox regression models: “age at baseline visit”, “sex”, “smoking history in pack-years”, “participant lives alone or lives with someone at baseline”, “Charlson co-morbidity index at baseline”, “UPDRS part 3 (motor) score”, “Hoehn & Yahr disease stage at baseline”, “Schwab & England ADL score at baseline”, and “Category of minimental state examination score at baseline”. And there are three categories of minimental state examination score at baseline. The first one is the minimental state examination score at baseline is less than 24, the second one is the score greater or equal to 24 and less than 28, and the last one is greater or equal than 28 and less than 30. 


## Outcome

The event of interest is institutionalisation, which is defined as entry to a nursing home, residential home, or long-stay hospital ward. Hence, the date of institutionalisation is the date of entry into the institution. And the censor is when patients who were lost to follow, who died without being institutionalised, or who were alive at end of follow-up but were not admitted to an institution. For participants who were not institutionalised, the censoring date will be the date lost to follow-up, if lost to follow-up; the date of death, if dead and not lost to follow-up; and date last known to be alive and living at home. The survival time is the difference between the event/censoring time and the date of baseline visit. In total, there are 575 participants, 315 patients with PD or atypical parkinsonism, and 140 events. 

## Statistical analysis

First, we used the Kaplan-Meier (KM) estimate to find the median time to institutionalization of participants with PD or Atypical parkinsonism and those in the control group. Separate KM survival curves were plotted and incidence rates were calculated for each of these three groups. 

Second, we used the data from patients with PD and atypical parkinsonism to identify the risk factors. Since the data was incomplete, the cluster of missing data and the pattern of missing data were plotted to see what variables tend to missing together and their pattern. The assumption of missing data was missing at random (MAR), hence we used the multiple imputation to impute patients’ missing data by predictive mean matching method with multivariate imputation by chained equations (MICE) algorithm. Since the patients' medical information of minimental state examination score at year 1 will not be used in the Cox regression model, we only use it as observed data to impute missing variables. There were 33 patients (10.5%) with missing medical records of minimental state examination score at baseline and smoking history in pack-years, hence, the imputation times was 10. Because the MICE algorithm is Markov chain Monte Carlo (MCMC) method, we plotted the parameters against the iteration number to determine whether the MICE algorithm has converged. Convergence was diagnosed as the plot showed no trend. The goodness-of-fit of the multiple imputation model were checked by the density plot of imputed versus observed data. Then, we built three Cox regression models to select predictors by comparing their heuristic shrinkage, $R^2$, Akaike information criterion (AIC) and Bayesian information criterion (BIC). The model without “UPDRS part 3 (motor) score” is the best model. The model assumption of proportional hazards was checked by Schoenfeld residuals method. The global p-value was larger than 0.05, hence the proportional hazards assumption was valid. We accessed the functional form of each continuous predictor by plotting martingale residual against continuous predictor. If the fitted model is correct, these residuals should symmetrically distributed with expected value 0. However, the plot showed trends in "age at baseline visit” and “Schwab & England ADL score at baseline”. To determine whether they need to be transformed or not, we compared models with and without restricted cubic splines (RCS) by $R^2$, AIC and BIC. The result showed that linear model fit was as good as a model with RCS. Hence, we chose the linear cox regression model. Finally, we presented the hazard ratios and confidence interval of estimate effects of predictors in model by figure and table. 

 


## Describe of data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,eval = T,warning = F,message = F)
```

```{r library}
library(xtable)
library(dplyr)
library(tidyr)
library(knitr)
library(gridExtra)
library(Hmisc)
library(readstata13)
library(survival)
library(survminer)
library(rms)
library(skimr)
library(finalfit)
library(mice)
library(ggplot2)
library(GGally)
library(png)
library(bookdown)
library(foreign)


# install.packages('Hmisc')
```

```{r options}
options( tinytex.verbose = TRUE)

```

```{r }
data<-read.dta13("updata3.dta",nonint.factors = T,missing.type = T)
data[is.na(data)] <- NA  #575
```


There are total `r nrow(data)` participants data,  with `r sum(is.na(data$dateinstitution))` participants' data undefined the date of institution.

```{r data}

attach(data)

data<-data %>%
  mutate(t=dateinstitution)%>%           
  mutate(t=coalesce(t,datelost))%>%
  mutate(t=coalesce(t,datedeath))%>%
  mutate(t=coalesce(t,datelastknownalive))

# sum(is.na(data$t))

data$cens<-ifelse(is.na(dateinstitution),0,1)


data$tt<-as.Date(as.character(data$t), format="%Y-%m-%d")-
                as.Date(as.character(data$datevisitbl), format="%Y-%m-%d")

data$tt<-as.numeric(data$tt)

data$year<-data$tt/365.25
```

```{r data2}

data2<- subset(data,!(diagnosiscat %in% c("Control")))
```



```{r eval=FALSE}

my_skim <- skim_with(base = sfl(n = length))

agebl.sum<-data%>%
  group_by(diagnosiscat)%>%
  dplyr::select(agebl)%>%
  my_skim()%>%
  transmute(Variable=skim_variable,Group=diagnosiscat,n=n,Mean=numeric.mean,SD=numeric.sd,Min=numeric.p0,Median=numeric.p50,Max=numeric.p100,IQR=numeric.p75-numeric.p25)




data$sex.mf<-ifelse(sex==0,"male","female")


sex.sum<-data%>%
  group_by(diagnosiscat,sex.mf)%>%
  dplyr::summarise(n=n())%>%
  mutate(percentage=paste0(round(100*n/sum(n),2),'%'))



my_skim <- skim_with(base = sfl(n = length))

packyears.sum<-data%>%
  group_by(diagnosiscat)%>%
  dplyr::select(packyears)%>%
  my_skim()%>%
    transmute(Variable=skim_variable,Group=diagnosiscat,Mean=numeric.mean,SD=numeric.sd,Median=numeric.p50,Max=numeric.p100,Q75=numeric.p75,Q25=numeric.p25)



data$eversmok.yn<-ifelse(eversmok==0,"no","yes")

eversmok<-data%>%
  group_by(diagnosiscat,eversmok.yn)%>%
  dplyr::summarise(n=n())%>%
  mutate(percentage=paste0(round(100*n/sum(n),1),'%'))%>%
  transmute(eversmok= eversmok.yn, n=n,percentage=percentage)
  

data$livesalone.yn<-ifelse(livesalone==1,"yes","no")

livesalone.sum<-data%>%
  group_by(diagnosiscat,livesalone.yn)%>%
  dplyr::summarise(n=n())%>%
  mutate(percentage=paste0(round(100*n/sum(n),1),'%'))


```

```{r datasum}

data.sum<-read.csv("Summary.csv")%>%
  rename("Characteristic"=ï..Characteristic,"Parkinson’s disease (N=201)"=Parkinsonâ..s.disease..N.201.,"Atypical parkinsonism (N=114)"=Atypical.parkinsonism..N.114.,"Control (N=260)"=Control..N.260.)

data.sum%>%
  kable(caption='\\label{tab:datasum} Summary information of participants.',booktabs=T,linesep="") 


```


Table \@ref(tab:datasum) shows there are 201 patients have been diagnosed Parkinson’s disease, and 114 patients have been diagnosed Atypical parkinsonism, and 260 patients in control group. The mean age at baseline visit for those with Parkinson’s disease is around 6 years younger than atypical parkinsonism. The youngest patient with Parkinson’s disease is 17.6 years younger than that from the atypical parkinsonism group. The interquartile range of patients’ baseline visit age with Parkinson’s disease is wider than the other two groups’ patients. And over 60% of all patients are male in each of the three groups. The widest interquartile range of smoking history in pack-years is in Atypical parkinsonism group. And the median of all three groups are much lower than their $75^{\mbox{th}}$ percentile values because the data are highly skewed, meaning that in each group there are lots of non-smoking participants. For example, more than half of PD patients don’t smoke. However, more than half of the patients smoke in the atypical parkinsonism and control groups. And there are some patients with missing data in the control group. Furthermore, the proportion of living alone in Atypical parkinsonism group (31.58%) is higher than that of Parkinson's disease (24.88%). And around half of the participants in control group live alone, despite there are three missing value in control groups. 


```{r ageboxplot, fig.height=4, fig.width=10, fig.align = "center",fig.cap = "\\label{ } Boxplots of patients' age in Parkinson’s disease and Atypical parkinsonism.", fig.pos ='H' }

plot1<-ggplot(data=data2,aes(y=agebl,x=diagnosiscat))+
  geom_boxplot()+
  ylab("Age at baseline visit")+
  xlab("Diagnoses group")

plot2<-ggplot(data=data2,aes(y=agedx,x=diagnosiscat))+
  geom_boxplot()+
  ylab("Age at diagnosis of parkinsonian syndrome")+
  xlab("Diagnoses group")


plot3<-ggplot(data=data2,aes(agesymptomonset,x=diagnosiscat))+
  geom_boxplot()+
  ylab("Age at onset of parkinsonian symptoms")+
  xlab("Diagnoses group")

grid.arrange(plot1, plot2, plot3,ncol=3, nrow=1)

```

The fig \@ref(fig:ageboxplot) shows the distribution of patients’ age at baseline, at diagnosis and at onset of parkinsonian symptoms are similarly. Hence, to avoid multicollinearity, only variable “age at baseline visit” will be included in the imputation and cox regression model. 




## Analysis


### The median time to institutionalization in each group.

```{r KM}

KM2<-survfit(Surv(tt/365.25,cens)~diagnosiscat,data=data)

data.frame(groups= c("Parkinson's disease","Atypical parkinsonism","Control"),n=c(201,114,260),events=c(71,69,45),median=c(11.78,2.55,NA),LCL=c(9.32,1.81,14.78),UCL=c(NA,4.22,NA))%>%
  kable(caption='\\label{tab:KM} Kaplan-Meier estimate of the survival function.',booktabs=T,linesep="",digits=2) 

```
Table \@ref(tab:KM) shows the median follow-up time for patients with Parkinson's disease is 11.78 years and for patients with Atypical parkinsonism is 2.55 years. The median survival time and the upper 95% confidence limit for control group are both undefined, indicated by a missing value “NA”. To better visualized the estimation, we plot the Kaplan-Maier survival curve.

### KM analysis of institutionalization with separate survival curves

```{r KMplot, fig.align = "center",fig.cap = "\\label { } Kaplan-Maier survival curve.", fig.pos ='H'}



ggsurvplot(KM2,xlab="Survival days",data,surv.median.line = "hv",legend="top",legend.labs=c("Parkinson's disease","Atypical parkinsonism","Control"))



```
The fig \@ref(fig:KMplot) shows that the survival curve for the Parkinson's disease group is above that for the Atypical parkinsonism group. Hence, we can conclude that the survival time for patients with Parkinson's disease is longer than patients with Atypical parkinsonism. And the control group has the longest survival time. 

### The incidence rate 

```{r incidence}

data%>%
  group_by(diagnosiscat)%>%
  summarise('incidence rate'=paste(round(sum(cens)/as.numeric(sum(year))*100,3),' per 100 person-years'))%>%
  rename(group=diagnosiscat)%>%
  kable(caption='\\label{tab:incidence} Summary of incidence rate by group.',booktabs=T,linesep="",digits=2) 



```

Table \@ref(tab:incidence) shows that Atypical parkinsonism group has the highest incidence rate of institutionalization (27.563 per 100 person-years), which is 5.3 times higher than the incidence rate in Parkinson's disease group (5.188 per 100 person-years). And the incidence rate in control group is the lowest, only 1.921%. 

### Identify the risk factors by cox regression model

```{r ,eval=FALSE}

#### Correlation of predictors

# ggcorr(data2[,c(6:18)], label=F)

# fig.align = "center",fig.cap = "\\label { } Corrleation plot.", fig.pos ='H'

# From the fig \@ref(fig:colplot), we can see the correlation between variable "agebl", "agedx", and "agesymptomonset" is one. Therefore, to avoid multicollinearity, only variable "agebl" will be included in the imputation model.
```


#### Missing data pattern

```{r missplot, fig.align = "center",fig.cap = "\\label { } Cluster analysis showing which predictors tend to be missing on the same patients.", fig.pos ='H'}

plot(naclus(data2[,c(6:18)]))

```

From the fig \@ref(fig:missplot), we can see that minimental state examination score at baseline and minimental state examination score at year 1 are likely to be found missing in the same patients. 

 
```{r mistable1}

explanatory<-c("packyears","mmsebltotal","mmseyr1total")
dependent<- c("cens","tt")

fit1<-data2%>%
  ff_glimpse(dependent,explanatory)

data.frame(Variable=c("smoking history in pack-years","minimental state examination score at baseline","minimental state examination score at year 1"),n=c(308,289,257),missing_n=c(7,26,58),missing_percent=c("2.2%","8.3%","18.4%")) %>%
  kable(caption='\\label{tab:mistable1}  Summary of missing data.',booktabs=T,linesep="",digits=1)

```

Table \@ref(tab:mistable1) shows that there are `r sum(is.na(data2$packyears))` missing in the variable of smoking history in pack-years. For minimental state examination score, `r sum(is.na(data2$mmsebltotal))` missing at baseline and `r sum(is.na(data2$mmseyr1total))` missing at year 1. 

```{r misspattern, fig.align = "center",fig.cap = "\\label { } Missing data pattern.", fig.pos ='H'}

mispattern<-data2 %>% 
  missing_pattern(explanatory)
```
The fig \@ref(fig:misspattern) is the visualisation of this missing pattern output, where the blue cubic means observed and red cubic means missing. From the fig \ref{fig:misspattern}, we can see that 26 patients missing with the minimental state examination score at baseline and 7 patients missing with smoking history in pack-years. The missing pattern is general and connected through the first row. The first column shows the frequency of each pattern and the number of missing entries in each pattern shows in the last column. For example, the first row means there are 241 patients in the dataset with all three observed variables. 



```{r ,eval=FALSE}

# as.data.frame(mispattern)%>%
#  rename(" "=V4)%>%
#  kable(caption='\\label{tab:mistable2}  Summary of missing data pattern.',booktabs=T,linesep="",digits=2)
```



#### Mutiple Imputation 


```{r pmm}

data2$diagnosiscatbinary<-ifelse(data2$diagnosiscat=="Atypical parkinsonism",0,1)   # (,yes,no) Atypical parkinsonism=0, Parkinson's disease=1

pmm<-mice(data2[,c(4,6,9:18,25:26)],method="pmm",seed=81420,m=10,print=F)


```




```{r convergence, fig.align = "center",fig.cap = "\\label{ } Mean and standard deviation of the synthetic values plotted against iteration number for the imputed data.", fig.pos ='H'}
# pmm$visitSequence

plot(pmm)

```
From fig \@ref(fig:convergence), we can see that the different streams freely intermingle with one another and there are no obvious trends. Hence, the MICE algorithm has converged.


```{r densityplot1,fig.height=5, fig.width=10, fig.align = "center",fig.cap = "\\label{ } Density plot of the observed data (blue) and the m=10 densities per variable calculated from the imputed data (thin red lines) of minimental state examination score at baseline ", fig.pos ='H' }

densityplot(pmm,~ mmsebltotal)
```

From the fig \ref{fig:densityplot1}, we can see that the distribution of observed and imputed data is similar.

```{r densityplot2, fig.height=5, fig.width=10, fig.align = "center",fig.cap = "\\label{ } Density plot of non-zero observed (blue) and the m=10 densities per variable calculated from the imputed data (thin red lines) of packyears.", fig.pos ='H'}

imp.packyears<-pmm[["imp"]][["packyears"]]


packyearsnozero<-subset(data2,packyears>0)


 plot(density(packyearsnozero$packyears),col="darkblue",main=" ", xlab = "packyears",xlim=c(-100,200),ylim=c(0.000,0.025))
for (i in 1:10) {
  lines(density(imp.packyears[,i] ),col="deeppink3")
}

```
In this study, the `r nrow(packyearsnozero)` patients with missing “packyears” variable are all smokers, which means pack years should be larger than zero. Therefore, fig \@ref(fig:densityplot2) shows the distribution for those with non-zero "packyears". We can see that the distribution between observed and imputed data is different.

The discrepancies may be caused by a bad imputation model, by a missing data mechanism that is not MCAR or by a combination of both. To further compare the distributions of observed and imputed data of "packyears", we use the diagnostic tool proposed by @Diagnostic2016. The idea is that under MAR the conditional distributions should be similar if the assumed model for creating multiple imputations has a good fit.

```{r diagnostic, fig.align = "center",fig.cap = "\\label{ } Packyears against missingness probability for observed and imputed values.", fig.pos ='H'}

fit<-with(pmm,glm(ici(pmm)~agebl+sex+packyears+eversmok+livesalone+charlsonbl+updrsblpart3+hybl+sebl+mmsebltotal+mmseyr1total+cens+tt,family = binomial))

# agebl+agesymptomonset+sex+packyears+eversmok+livesalone+charlsonbl+updrsblpart3+hybl+sebl+mmsebltotal+mmseyr1total+cens+tt

ps<-rep(rowMeans(sapply(fit$analyses,fitted.values)),pmm$m +1)

xyplot(pmm, packyears~ps | as.factor(.imp),
       xlab="Probability that record is incomplete",
       ylab="Packyears",
       pch=c(1,19),
       col=mdc(1:2))




```

The fig \@ref(fig:diagnostic) shows that the distributions of the blue and red points are quite similar, as expected under MAR. Therefore, the imputation model has a good fit.



```{r mmsecat}

long<-complete(pmm, 'long', include=TRUE) #include=T means, the long contains original data and 10 imputed data.
                                          #if include=FALSE or not specified, Error in as.mids(long) : Original data not found.


N <- nrow(long)
for (i in 1:N){
  if(!is.na(long$mmsebltotal[i])){
    if (long$mmsebltotal[i] <= 23 ){
      long$mmsecat[i] = 0
    }else if (long$mmsebltotal[i] <= 27 && long$mmsebltotal[i] > 23){
      long$mmsecat[i] = 1
    }else if (long$mmsebltotal[i] <= 30 && long$mmsebltotal[i] > 27){
      long$mmsecat[i] = 2
    }
  }
}




long$mmsecat<-factor(long$mmsecat,
                        levels = c(0,1,2),
                        labels = c("<24", "24-27", "28-30"))



pmm.new<-as.mids(long)



```







#### Model selection

Since the UPDRS part 3 (motor) score and Hoehn & Yahr disease stage are both used to evaluate the degree of disability and measure the progress of Parkinson’s symptoms, hence they are similar predictors. To determine whether they should both include in the model or not, we compare 3 Cox regression models to select predictors. The first is the full model with all these nine predictors, the second is the model without “hybl” variable and the third is the model without “updrsblpart3”.

```{r cox models,results = 'hide'}

dd<-datadist(long)
options(datadist='dd')

modela<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+hybl+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new)  #full

modelb<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) #without hybl

modelc<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) #without updrsblpart3

```

```{r ,eval=FALSE}
modela
modelb
modelc

```



```{r model select}

AIC1<-AIC(modela)  #full
AIC2<-AIC(modelb)  #without hybl
AIC3<-AIC(modelc)  #without updrs


BIC1<-BIC(modela)  #full
BIC2<-BIC(modelb)  #without hybl
BIC3<-BIC(modelc) #without updrs


data.frame(Model=c("Full Model","Model without Hoehn & Yahr disease stage at baseline","Model without UPDRS part 3 (motor) score") , 'LR chi2'=c(149.81,147.68,149.76),'d.f.'=c(10,9,9),'R square' = c(0.383,0.379,0.383),AIC = c(AIC1,AIC2,AIC3), BIC = c(BIC1,BIC2,BIC3))%>%
  kable(caption='\\label{tab:model select} Model tests and discrimination indexes.',booktabs=T,linesep="",digits=2)


```


Table \@ref(tab:model select) shows that the model without UPDRS part 3 (motor) score has highest $R^2$ and the lowest AIC and BIC value. And the heuristic shrinkage estimate for the full model is 0.93 (139.81/149.81), for the model without Hoehn & Yahr disease stage at baseline is 0.94 (138.68/147.68), and for the model without UPDRS part 3 (motor) score is 0.94 (140.76/149.76). Therefore, these three models are all fitted the data well and the noise is very low. Due to this study only has 140 events and to avoid over fitting, we choose the model without UPDRS part 3 (motor) score.


Then we test the proportional hazards assumption of this cox regression model by Schoenfeld residuals in each of the imputed datasets.

```{r testcox}


pmm.new_names <- paste("pmm.new",1:10,sep="")  # create names first

for (i in 1:10){
    pmm.new_df<-complete(pmm.new,i)
    assign(pmm.new_names[i], pmm.new_df)
}


test_names<- paste("test.ph",1:10,sep="")   # create names first

for (i in 1:10){
    test.ph_df<-cox.zph(coxph(Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat,data=get(paste0("pmm.new",i))),transform = "km")   # data=get(paste0("pmm.new",i))
    assign(test_names[i], test.ph_df)
}

unlist(test.ph10[["table"]]) %>%
   kable(caption='\\label{tab:testcox} PH assumption test results.',booktabs=T,linesep="",digits=3)
```

Table \@ref(tab:testcox) shows a example of proportional hazards assumption, we can see that except for "Schwab & England ADL score at baseline", all predictors' p-value is larger than 0.05. It indicates that only the "Schwab & England ADL score at baseline" effect changes significantly over time. However, the p-value of global test of PH assumption is larger than 0.05. Hence there is not enough evidence to reject the null hypothesis of proportional hazards. After checking proportional hazards assumption in each imputed dataset, we found no evidence to reject the assumption.


```{r,eval=FALSE}
plot(cox.zph(coxph(Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat,data=pmm.new10),transform = "km"),var="sebl")


# fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat+sebl:tt),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new)



```

```{r,eval=FALSE}

#page 545

z<-predict(modelc,type = 'terms')
z.sebl<-z[,7]
z.other<-z[,-7]
f.ia<-cph(Surv(data2$tt,data2$cens)~z.sebl*z.other)
anova(f.ia)

```








```{r martingale,fig.align = "center",fig.cap = "\\label{ } Martingale residuals versus four continuous predictors.", fig.pos ='H'}

pmm.new1<-complete(pmm.new,1)
rr.0<-residuals(coxph(Surv(tt,cens)~1,data = data2),type="martingale")

pmm.new1$rr.0<-rr.0


plot1<-ggplot(pmm.new1,aes(agebl, rr.0)) + 
         geom_point(cex=1) +
         geom_smooth(method = "loess", se = T)+
         ylim(c(-2,2))



plot2<- ggplot(pmm.new1,aes(packyears, rr.0)) + 
         geom_point(cex=1) +
         geom_smooth(method = "loess", se = T) +
         ylim(c(-2,2))   


plot3<- ggplot(pmm.new1,aes(sebl, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        ylim(c(-2,2))

plot4<-ggplot(pmm.new1,aes(hybl, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        ylim(c(-2,2))

grid.arrange(plot1,plot2,plot3,plot4, ncol=1, nrow=4)


```
From the \@ref(fig:martingale) we can see that "smoking history in pack-years" seems to be symmetrically distributed around $\pm 1$ and the blue fitted lowess curve seems straight and flat. The scatter points of "Hoehn & Yahr disease stage at baseline" seems symmetrically distributed around $\pm 0.5$ with expected value of zero. For the scatter plot of "Schwab & England ADL score at baseline", there appears to be a slightly decreasing trend  while for "age at baseline visit" the line is slightly going up. 


To determine whether "age at baseline visit" and "Schwab & England ADL score at baseline" need to be transformed or not, we compare models with and without restricted cubic splines (RCS).

```{r rcs,results = 'hide'}

modelc1<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age as 3 knot rcs


modelc2<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+rcs(sebl,3)+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with sebl as 3 knot rcs


modelc3<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+hybl+rcs(sebl,3)+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age and sebl as 3 knot rcs

```


```{r ,eval=FALSE}
modelc1
modelc2
modelc3

```


```{r model select2}
AIC.C<-AIC(modelc)    # chose one,without updrs
AIC.C1<-AIC(modelc1)  # with age as 3 knot rcs
AIC.C2<-AIC(modelc2)  # with sebl as 3 knot rcs
AIC.C3<-AIC(modelc3)  # with age and sebl as 3 knot rcs

BIC.C<-BIC(modelc)   # chose one,without updrs
BIC.C1<-BIC(modelc1)  # with age as 3 knot rcs
BIC.C2<-BIC(modelc2)  # with sebl as 3 knot rcs
BIC.C3<-BIC(modelc3)  # with age and sebl as 3 knot rcs




data.frame(Model=c("Model without RCS","Model with age 3 knots RCS","Model with sebl 3 knots RCS","Model with age and sebl 3 knots RCS") , 'LR chi2'=c(149.76,149.81,152.00,152.06),'d.f.'=c(9,10,10,11),'R square' = c(0.383,0.383,0.387,0.387),AIC = c(AIC.C,AIC.C1,AIC.C2,AIC.C3), BIC = c(BIC.C,BIC.C1,BIC.C2,BIC.C3))%>%
  kable(caption='\\label{tab:model select2} Model with and without RCS tests and discrimination indexes.',booktabs=T,linesep="",digits=2)


```
Table \@ref(tab:model select2) shows that there is very little difference between the model $R^2$, AIC and BIC value of models with or without RCS terms. The model with sebl 3 knots RCS has the highest $R^2$ and lowest AIC value. However, the $R^2$ value of the linear model is only 0.04 lower than the highest one and the AIC is only 0.293 higher than the lowest one. The linear model has the lowest BIC. Therefore, since the simplest model is just as good as the models with RCS terms, we choose the model without RCS.


Then we present the point and interval estimates of predictor effects by a hazard ratio chart and a table of summary estimated effects. 
```{r coxoutplot,out.width="70%",fig.align = "center",fig.cap = "\\label{ } Hazard ratios and confidence bars for effects of predictors in model.", fig.pos ='H'}

plot(summary (modelc),main= '',log = T)


```

```{r coxout}

sumrcs2<-data.frame(summary(modelc))

sumrcs2.new<-sumrcs2[grep("Hazard.Ratio",rownames(sumrcs2)),]

row.names(sumrcs2.new)<-c("agebl","sex","packyears","livesalone","charlsonbl","hybl","sebl","mmsecat - <24:28-30 ","mmsecat - 24-27:28-30")


sumrcs2.new%>%
  select(Effect,Lower.0.95,Upper.0.95)%>%
  rename("Hazard ratio"=Effect)%>%
   kable(caption='\\label{tab:coxout} Summary of estimated effects.',booktabs=T,linesep="",digits=1)
  

```

Table \@ref(tab:coxout) shows the hazard ratio confidence interval of variable "sex", "livesalone", "charlsonbl", "hybl", and "sebl" include one. Hence, they are not statistical significance. The hazard of being admitted into institution for female is 11.8% higher than male. And for patients live alone the hazard is 22.1% higher than those live with someone at baseline. And patients with one higher level in Hoehn & Yahr disease stage at baseline will increase the hazard by 29.9%. However, patients with one level higher in Charlson co-morbidity index or Schwab & England ADL score at baseline, their hazard ratio of being admitted into institution will decrease 84.4% and 71.1% respectively. 


On the contrary, the effects of "agebl", "packyears" and "mmsecat" are statistical significance. The hazard ratio of patients who are one year old older than others at baseline visit is 2.01. Hence, every one year older, the hazard of being admitted into institution increases 100.5%. And for patients'  one more pack in annual smoking history, the hazard of being admitted into institution increases 18.3%. And patients with higher minimental state examination score at baseline are tend to have less hazard. Patients with minimental state examination score at baseline less than 24 will have 211.7% more hazard of entering institution than those with score at baseline equal or larger than 28 and less than 30, and those score between 24 to 27 will have 85.3% more hazard than those with score equal or larger than 28 and less than 30. 

```{r nomogram, eval=FALSE}



surv <- Survival(modelc)
surv5 <- function(x) surv(5*365,lp=x)
surv8 <- function(x) surv(8*365,lp=x)


plot(nomogram(modelc, 
              fun=list(surv5, surv8), 
              funlabel=c('5-year survival', '8-year survival')))

```



```{r, eval=FALSE}


img1_path<-"nomogram.png"
img1<-readPNG(img1_path,native = T,info = T)
# attr(img1,"info")
include_graphics(img1_path)
```





## Summary

For same years old patients, the risk of woman being admitted into institution is higher than male. And if the patient's age is above median age at baseline visit and living alone, clinicians should pay more attention to him/her need for help with daily activities and plan management earlier. patients' smoking habit will also increase their chance into nursing home. Patients with heavy smoking habit should be advised to live a more health life and smoking less. In general, the highest risk patient's profile is older than 80 years old heavy smoker woman who lives alone.

If two patients' personal profile is the same, then their medical profile will effect their chance of entering nursing home. Hoehn and Yahr scale is a common way to describe the Parkinson's disease progress. The higher Hoehn and Yahr disease stage at baseline means the less independent. To assess patients' impaired mobility, we choose the the Schwab and England ADL score which represents the percentage of independence. The lower the score is, the more dependent of patients and hence more care will be needed. Some patients' may not suffer from functional disability in body, but suffer from cognitive impairment. The Mini-Mental State Exam (MMSE) is to test patients' mental skills. The lower mini-mental state examination score at baseline means the more severe dementia in patient. And compare to impaired mobility, the cognitive impairment affect patients' independence more seriously. However, sometimes patients may die before being admitted into institution. The Charlson co-morbidity index is used to predict patients' one-year mortality. Therefore, patients with higher Charlson co-morbidity index at baseline will decrease the chance of being admitted into institution due to their living time is shorter than waiting time. In short, patients with higher Hoehn and Yahr scale and lower in Schwab and England ADL score and mini-mental state examination score at baseline will have higher demand of dependency and will need clinicians to plan management earlier. Meanwhile, patients with high Charlson co-morbidity index should be priority consider to entering nursing home. 


## Strengths and limitations

There are only 315 patients be involved in this study and only 140 events. Therefore, the sample size of this study is not big enough and will arise some bias. Due to the sample size problem, we have to choose the linear cox regression model to avoid over fitting. Therefore, the confidence interval might be wider. The sample size is also limited the number of predictors. And in the cox regression model, the PH assumption is valid in global test but not in "sebl" variable. 

## Further work 

In further analysis, it could expand the patients' number in the study and use more patients' profile personal information as predictors to build cox regression model. And this model can be validated in external data. 




# Reference


