---
title: "Dynmatic predtion model in parkinson disease"
author: "Yan Li"
header-includes:
- \usepackage{floatrow}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \floatsetup[table]{capposition=top}
- \floatplacement{table}{H}
output:
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
        scroll_highlight: yes
      before: null
      after: null
    toolbar:
      position: fixed
    edit : null
    download: [["book.pdf", "Download PDF version"]]
    search: yes
    fontsettings:
      theme: white
      family: sans
      size: 2
    sharing:
      facebook: yes
      github: no
      twitter: yes
      linkedin: no
      weibo: no
      instapaper: no
      vk: no
      whatsapp: no
      all: ['facebook', 'twitter', 'linkedin', 'weibo', 'instapaper']
    info: yes
  bookdown::html_book:
    css: toc.css
  bookdown::pdf_book:
    keep_tex: yes
documentclass: book
bibliography: my5.bib
fontsize: 14pt
citation_package: none
csl: harvard-cite-them-right2.csl
nocite: '@*'
---
# Systematic review of prognostic models in Parkinson’s disease

# PD cox regression

## Abstract

When patients with Parkinson’s disease lose their ability to take care of themselves, and carers are no longer able to look after them at home, they will need long-term care. The risk factors for being admitted to a nursing home are unclear and there are few published studies. The aim of this work is to identify the incidence of institutionalization and the risk factors for institutionalization using data from a prospective incident cohort of parkinsonism in North East Scotland. This prospective cohort study has collected annual follow-up data from 575 participants from Aberdeen, Scotland, since 2002.  


with the baseline visit time from 2002 to 2010.


Using the reverse Kaplan-Meier method, the median time to institutionalization of patients with Parkinson’s disease (11.8 years) was 4.7 times longer than Atypical parkinsonism (2.5 years). The results from the Cox regression showed that patients with mild dementia (versus )are at highest risk for entering institutionalization (HR=3.1) and ageing increasing the need of the long-term care (HR=2). While the sex of patients (HR=1.1), smoking habits (HR=1.2) and whether living alone (HR=1.2) are all not high-risk factors, since their 95% confidence interval all contain one. Therefore, older patients with mild dementia, especially Atypical parkinsonism patients, are more likely need long-term care.







## Introduction



Parkinson’s disease (PD) and atypical parkinsonism are neurodegenerative disorders. The cause of PD is unknown and both the genetic and environmental factors affect it. Moreover, the diagnosis of PD is hard to make and there are no treatment can slow the neurodegenerative process. As the disease progresses, the need for care in PD patients will increase. In the late stage, the dopaminergic therapies are lack of efficiency to the motor symptoms (dysphagia, postural instability, freezing of gait and falls), and the common non-motor features appear (urinary symptoms, orthostatic hypotension, and dementia). The motor and non-motor features in PD patients are heterogeneous, but they all reduce patients' quality of life and the independent living ability. Hence, many patients will need for enter a long-term care institution, such as a nursing home. However, the management strategies for PD patients in late stage are poor [@Lancet2015]. Several studies have investigated risk factors for institutionalisation. The first study was a retrospective case-control study with only 11 PD patients and found hallucinations was the risk factor for nursing home placement [@pred1993]. Then @pred2000 carried out a four-year prospective study (population-based study in western Norway), which followed 178 community-dwelling subjects in PD from baseline to nursing home admission, and found the highest risk were hallucinations, cognitive impairment and old age. A nationwide retrospective cohort study followed 469055 PD elderly patients from 2002 to December 2005, and found woman and black individuals with PD were more likely entering nursing facility residence compared with men and white individuals [@pred2015]. However, no previous studies have reported the incidence of institutionalisation in PD and atypical parkinsonism in a population-representative cohort. No previous studies have investigated risk factors in a population-representative cohort in parkinsonism. This study will use data from Parkinsonism incidence in North-East Scotland (PINE) study. We will first identify the incidence of institutionalization in PD and atypical parkinsonism. Then identify the risk factors for institutionalization in PD and atypical parkinsonism. The results from this study will enable clinicians to manage their patients better and earlier and provide more appropriate services to help PD patients.


## Methods

**Study population,design and clinical characteristics**

 
The Parkinsonism incidence in North-East Scotland (PINE) study is a prospective and population-based incident cohort that follows all newly diagnosed patients with degenerative or vascular parkinsonism along with age-sex matched control group and collects annual patient data on survival, disability, and institutionalization from 2002 to 2020 around Aberdeen, Scotland [@Field2016]. The patients' information updates each year and for those who died, the details about death are collected from death certificates and primary care and secondary care records. In this study, there are 201 PD patients, 114 Atypical parkinsonism patients and 260 control individuals. The baseline visit time is from November 2002 to June 2010, and the last follow up time is April 2020. The average follow-up time of each participant is 6.9 years. And 185 individuals entered (45 in control group) institutionalization, 420 death (142 in control group) and 6 lost follow up (3 in PD and 3 in control group). 

Parkinson’s rating scales are used to assess patients' symptoms of the condition and evaluate treatment, and clinicians assign a value to it depends on patients' particular feature information. The Unified Parkinson’s Disease Rating Scale (UPDRS) has four parts and the part three is motor examination and higher score means more server (0-4 rating scale). And the UPDRS is often used with the Hoehn and Yahr Scale and the Schwab and England Activities of Daily Living (ADL) Scale [@Ratingscales]. The Hoehn and Yahr Scale is used to measure the progress of Parkinson’s symptoms and the level of disability. It includes stages 0 to 5, where stage 0 means no signs of disease and stage 5 means need a wheelchair or bedridden. The Schwab and England ADL Scale is to measure patients' independence of daily activities through a percentage figure, the high percentages indicate independence while low percentages indicate dependence. Patients' death risk and mental abilities were also measured. Charlson co-morbidity index categorizes patients' comorbid conditions by assigned a weight from 1 to 6 to each condition and it includes 16 conditions [@Charlson]. The Mini Mental State Examination (MMSE) test is used to diagnose dementia and assess its progression and give indication of severity. The maximum score of MMSE is 30 points, less than 24 points indicates mild dementia. And the MMSE score of Alzheimer's patients drops around two to four points averagely each year [@MMSE]. Hence, if patient's medical history record without the baseline MMSE score, we can use MMSE score on year one to infer the MMSE score on baseline. 

The annual clinical examination includes: (i) parkinsonian impairment (UPDRS part 3); (ii) disease stage (Hoehn and Yahr Scale); (iii) disability (Schwab and England Activities of Daily Living (ADL) Scale); (iv) comorbid conditions (Charlson co-morbidity index); (v) cognitive function (Mini Mental State Examination (MMSE) test). The demographic information of patients in baseline visit includes: (i) age; (ii) sex; (iii) smoking history in pack-years; (v) whether lives alone or not. There are 7 individuals missing the data of smoking history in pack-years (2.2%), 26 individuals missing the data of MMSE test score at the baseline (8.3%) and 58 individuals missing the data of MMSE test score at year one (18.4%). 


```{r eval=FALSE,echo=FALSE}

#data$n.death<-ifelse(!is.na(datedeath),1,0)

#data%>%
#  group_by(diagnosiscat)%>%
#  summarise(sum=sum(n.death))
  

#data$n.lost<-ifelse(!is.na(data$datelost),1,0)

#data%>%
#  group_by(diagnosiscat)%>%
#  summarise(sum=sum(n.lost))
 
```


**Outcome**

The event of interest is institutionalisation, which is defined as entry to a nursing home, residential home, or long-stay hospital ward. Hence, the date of institutionalisation is the date of entry into the institution. And the censor is when patients who were lost to follow, who died without being institutionalised, or who were alive at end of follow-up but were not admitted to an institution. For participants who were not institutionalised, the censoring date will be the date lost to follow-up, if lost to follow-up; the date of death, if dead and not lost to follow-up; and date last known to be alive and living at home. The survival time is the difference between the event/censoring time and the date of baseline visit. In total, there are 575 participants, 315 patients with PD or atypical parkinsonism, and 140 events. 

**Statistical analysis**


The Kaplan-Meier (KM) estimate were plotted to find the median time to institutionalization of participants with PD or Atypical parkinsonism and those in the control group. KM estimate is an nonparametric estimation of cumulative survival probability, it gives the survival probability of every unique event time. At each time of the event happens, the cumulative survival probability is calculated by one minus the number of individuals who enter the institutionalisation divides the number of patients at risk. Therefore, at the begin of time zero, the cumulative survival probability is always one and the survival probability will change over time. At the time of event, the plot of KM estimate will have a "step". Hence, more events happen, the plot will be more smooth. Separate KM survival curves were plotted and incidence rates were calculated for each of these three groups. The plot is survival probability against the time, we can see how the survival probability of these three groups change over time and see how different/similar they are. 



To identify the risk factors in institutionalization, we used the data excludes the control group. Since the data was incomplete, the cluster of missing data and the pattern of missing data were plotted to see what variables tend to missing together and their pattern. The assumption of missing data was missing at random (MAR), hence we used the multiple imputation to impute patients’ missing data by predictive mean matching method with multivariate imputation by chained equations (MICE) algorithm. The multiple imputation is to use the existing information of the data to impute the missing data and the predictive mean matching method do not need to require data distribution. For example, patients who miss the Mini Mental State Examination (MMSE) test score at baseline can be imputed by the data of MMSE test at year one and other existing information. The value that need to be impute is follow the distribution of the the posterior probability of missing data given the observed data and it repeats imputation times drawing data from the distribution. There were 33 patients (10.5%) missing with MMSE test score at baseline and smoking history in pack-years, hence, the imputation times was 10. Because the MICE algorithm is Markov chain Monte Carlo (MCMC) method, we plotted the parameters against the iteration number to determine whether the MICE algorithm has converged. Convergence was diagnosed as the plot showed no trend. The goodness-of-fit of the multiple imputation model were checked by the density plot of imputed versus observed data. 


Cox regressions were built to identify the hazard ratio of potential risk factors. Cox regression is a semiparametric model for a time-to-event outcome and it makes no assumption of the data. The model is composed by baseline hazard ratio and relative hazard function, $h(t|X)=h_{0}(t)exp(X\beta)$. The relative hazard function is an exponential formula of risk score which combines characteristics of an individual. The assumption is proportional hazards, which means the hazard ratio of predictors does not change over time. But it dose not assume the hazard function is constant. The Cox regression can give the hazard ratio of predictors, which is the ratio of hazard rates between two groups and higher hazard ratio means more risk. 

Due to the UPDRS part 3 and Hoehn and Yahr Scale are both Parkinson’s rating scales that used to assess patients' symptoms, they may have similar predictive effect. Therefore, we first built three Cox regression models: (i) Full model; (ii) Model without Hoehn and Yahr Scale; (iii) Model without UPDRS part 3. Then, we compared these three models by heuristic shrinkage, $R^2$, Akaike information criterion (AIC) and Bayesian information criterion (BIC) to select predictors. Shrinkage shows the difference between real value and the predict value, and the value is from zero to one. Higher shrinkage value means the model more closed to the real model. The heuristic shrinkage is calculated by one minus the model chi-square divides by the degrees of freedom. $R^2$ gives the proportion of the variability of the dependent variable that can be explained by the independent variable. It is used to measure the goodness of it of the model and ranges from zero to one. $R^2$ of one indicates model is perfectly fit and it's equation is one minus the sum of squares of residuals divides by the total sum of squares. AIC also measure the goodness of fit of model and the lower AIC value means the model fits better. BIC is used to compare model with different numbers of parameters, model with lower BIC value is more likely to be the true model. AIC tends to choose a bigger model, while BIC tends to choose a small model because BIC penalizes model complexity more heavily. Hence, AIC and BIC are used together to choose model. The formal is $-2logL+kp$, where L is the likelihood function and p is the number of parameters in the model, and k is 2 for AIC and log(n) for BIC.  

After selecting the model, the model assumption of proportional hazards was checked by Schoenfeld residuals method. The Schoenfeld residual is the observed value for patient minus the expected value of this patient. 
And under the proportional hazards assumption, plot of the scale Schoenfeld residuals of each predictors over time should not have trends. If the Schoenfeld residual plot can fit a straight line, then the p-value of the hypothesis test of proportional hazards assumption should be larger than 0.05. Then, we accessed the functional form of each continuous predictor by plotting martingale residual against continuous predictor. If the fitted model is correct, these residuals should symmetrically distributed with expected value 0. However, the plot showed trends in "Age” and “Schwab & England ADL Scale”. To determine whether they need to be transformed or not, we compared models with and without restricted cubic splines (RCS) by $R^2$, AIC and BIC. Finally, we presented the hazard ratios and confidence interval of estimate effects of predictors in model by figure and table. 

 


## Results



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,eval = T,warning = F,message = F)
```

```{r library}
library(xtable)
library(dplyr)
library(tidyr)
library(knitr)
library(gridExtra)
library(Hmisc)
library(readstata13)
library(survival)
library(survminer)
library(rms)
library(skimr)
library(finalfit)
library(mice)
library(ggplot2)
library(GGally)
library(png)
library(bookdown)
library(foreign)
library(kableExtra)

# install.packages('Hmisc')
```

```{r options}
options( tinytex.verbose = TRUE)

```

```{r }
data<-read.dta13("updata3.dta",nonint.factors = T,missing.type = T)
data[is.na(data)] <- NA  #575
```

```{r data}

attach(data)

data<-data %>%
  mutate(t=dateinstitution)%>%           
  mutate(t=coalesce(t,datelost))%>%
  mutate(t=coalesce(t,datedeath))%>%
  mutate(t=coalesce(t,datelastknownalive))

# sum(is.na(data$t))

data$cens<-ifelse(is.na(dateinstitution),0,1)


data$tt<-as.Date(as.character(data$t), format="%Y-%m-%d")-
                as.Date(as.character(data$datevisitbl), format="%Y-%m-%d")

data$tt<-as.numeric(data$tt)

data$year<-data$tt/365.25


# str(data)

data$sex<-factor(data$sex,
                        levels = c(0,1),
                        labels = c("male", "female"))

data$livesalone<-factor(data$livesalone,
                        levels = c(0,1),
                        labels = c("no", "yes"))
```

```{r data2}

data2<- subset(data,!(diagnosiscat %in% c("Control")))
```

```{r eval=FALSE}

my_skim <- skim_with(base = sfl(n = length))

agebl.sum<-data%>%
  group_by(diagnosiscat)%>%
  dplyr::select(agebl)%>%
  my_skim()%>%
  transmute(Variable=skim_variable,Group=diagnosiscat,n=n,Mean=numeric.mean,SD=numeric.sd,Min=numeric.p0,Median=numeric.p50,Max=numeric.p100,IQR=numeric.p75-numeric.p25)


data$sex.mf<-ifelse(sex==0,"male","female")


sex.sum<-data%>%
  group_by(diagnosiscat,sex.mf)%>%
  dplyr::summarise(n=n())%>%
  mutate(percentage=paste0(round(100*n/sum(n),2),'%'))



my_skim <- skim_with(base = sfl(n = length))

packyears.sum<-data%>%
  group_by(diagnosiscat)%>%
  dplyr::select(packyears)%>%
  my_skim()%>%
    transmute(Variable=skim_variable,Group=diagnosiscat,Mean=numeric.mean,SD=numeric.sd,Median=numeric.p50,Max=numeric.p100,Q75=numeric.p75,Q25=numeric.p25)



data$eversmok.yn<-ifelse(eversmok==0,"no","yes")

eversmok<-data%>%
  group_by(diagnosiscat,eversmok.yn)%>%
  dplyr::summarise(n=n())%>%
  mutate(percentage=paste0(round(100*n/sum(n),1),'%'))%>%
  transmute(eversmok= eversmok.yn, n=n,percentage=percentage)
  

data$livesalone.yn<-ifelse(livesalone==1,"yes","no")

livesalone.sum<-data%>%
  group_by(diagnosiscat,livesalone.yn)%>%
  dplyr::summarise(n=n())%>%
  mutate(percentage=paste0(round(100*n/sum(n),1),'%'))


```

**Describe of data**

```{r datasum}

data.sum<-read.csv("demographic summary.csv")%>%
  rename("Characteristic"=ï..Characteristic,"Parkinson’s disease (N=201)"=Parkinsonâ..s.disease..N.201.,"Atypical parkinsonism (N=114)"=Atypical.parkinsonism..N.114.,"Control (N=260)"=Control..N.260.)

data.sum%>%
  kable(caption='\\label{tab:datasum} Summary demographic information of participants.',booktabs=T,linesep="") 


```

Table \@ref(tab:datasum) shows the population in this study is elderly participants as most of them are over 70 years old and the sex ratio between male and female are around 1.5. Both age and sex distributions are similar in these three groups. There are more individuals have smoking habits and living alone in the control group, while more heavy smoker in atypical parkinsonism group. And less living alone patients in PD group than others. In short, the demographic information of participants are very similar among these three group with slightly difference in smoking habit and whether living alone. 




```{r  eval=FALSE}

#ageboxplot, fig.height=4, fig.width=10, fig.align = "center",fig.cap = "\\label{ } Boxplots of patients' age in Parkinson’s disease and Atypical parkinsonism.", fig.pos ='H'

plot1<-ggplot(data=data2,aes(y=agebl,x=diagnosiscat))+
  geom_boxplot()+
  ylab("Age at baseline visit")+
  xlab("Diagnoses group")

plot2<-ggplot(data=data2,aes(y=agedx,x=diagnosiscat))+
  geom_boxplot()+
  ylab("Age at diagnosis of parkinsonian syndrome")+
  xlab("Diagnoses group")


plot3<-ggplot(data=data2,aes(agesymptomonset,x=diagnosiscat))+
  geom_boxplot()+
  ylab("Age at onset of parkinsonian symptoms")+
  xlab("Diagnoses group")

grid.arrange(plot1, plot2, plot3,ncol=3, nrow=1)

```

 
**Kaplan-Meier estimation** 


```{r reKM}

data$cens.followup<-1-data$cens

reverse.KM<-survfit(Surv(tt/365.25,cens.followup)~diagnosiscat,data=data)

# KM2

data.frame(groups= c("Parkinson's disease","Atypical parkinsonism","Control"),n=c(201,114,260),median=c(9.26,4.41,10.78),LCL=c(8.30,3.33,10.40),UCL=c(11.80,5.78,11.07))%>%
  kable(caption='Reverse Kaplan-Meier estimate of the median follow-up time.',booktabs=T,linesep="",digits=2) %>%
  footnote(alphabet = c ("The time is year",
               "LCL is the 95% lower confidence limits;", 
               "UCL is the 95% upper confidence limits;"
               ))
  

```
Table \@ref(tab:reKM) shows the median follow-up time for PD patients (9.26 years) is 2.1 times longer than atypical parkinsonism patients (4.41 years). 


```{r KMplot, fig.align = "center",fig.cap = "\\label { } Kaplan-Maier survival curve.", fig.pos ='H'}

KM<-survfit(Surv(tt/365.25,cens)~diagnosiscat,data=data)

ggsurvplot(KM,ylab="Survival probability of institutionalisation",xlab="Survival years",data,legend="top",surv.median.line = "hv",legend.labs=c("Parkinson's disease","Atypical parkinsonism","Control"))


```
The fig \@ref(fig:KMplot) shows that at the beginning the survival probability of entering the institutionalisation is overlap between PD group and control group, however the survival probability for atypical parkinsonism group drops very quick. And when the follow time around 5 years, the atypical parkinsonism group's KM curve is not that smooth as before, which means most patients with atypical parkinsonism have already been in institutionalisation. The KM curve for PD patients is smooth and the survival probability is higher than atypical parkinsonism group all the time. Hence, patients with PD have longer independent living time than patients with atypical parkinsonism. And the median survival time for PD patients entering long-term care facility (11.8 years) is 4.7 times longer than atypical parkinsonism patients (2.5 years).


**The incidence rate**

```{r incidence}



data%>%
  group_by(diagnosiscat)%>%
  summarise('Incidence rate'=paste(round(sum(cens)/as.numeric(sum(year))*100,1),' per 100 person-years'))%>%
  rename(Group=diagnosiscat)%>%
  kable(caption=' The incidence rate of institutionalisation.',booktabs=T,linesep="") 



```

Table \@ref(tab:incidence) shows that atypical parkinsonism group (27.6 per 100 person-years) has 5.3 times higher incidence rate of institutionalisation than PD group (5.2 per 100 person-years). Therefore, atypical parkinsonism patients have higher risk of entering institutionalization than PD patients.


**Multiple imputation**



```{r ,eval=FALSE}

#### Correlation of predictors

# ggcorr(data2[,c(6:18)], label=F)

# fig.align = "center",fig.cap = "\\label { } Corrleation plot.", fig.pos ='H'

# From the fig \@ref(fig:colplot), we can see the correlation between variable "agebl", "agedx", and "agesymptomonset" is one. Therefore, to avoid multicollinearity, only variable "agebl" will be included in the imputation model.
```




```{r missplot, fig.align = "center",fig.cap = "\\label { } Cluster analysis showing which predictors tend to be missing on the same patients.", fig.pos ='H'}

plot(naclus(data2[,c(6:18)]))

```

From the fig \@ref(fig:missplot), we can see that Mini Mental State Examination test score at baseline and Mini Mental State Examination test score at year 1 are likely to be found missing in the same patients. 

 
```{r eval=FALSE}

explanatory<-c("packyears","mmsebltotal","mmseyr1total")
dependent<- c("cens","tt")

fit1<-data2%>%
  ff_glimpse(dependent,explanatory)

data.frame(Variable=c("smoking history in pack-years","minimental state examination score at baseline","minimental state examination score at year 1"),n=c(308,289,257),missing_n=c(7,26,58),missing_percent=c("2.2%","8.3%","18.4%")) %>%
  kable(caption='\\label{tab:mistable1}  Summary of missing data.',booktabs=T,linesep="",digits=1)


#Table \@ref(tab:mistable1) shows that there are `r sum(is.na(data2$packyears))` missing in the variable of smoking history in pack-years. For minimental state examination score, `r sum(is.na(data2$mmsebltotal))` missing at baseline and `r sum(is.na(data2$mmseyr1total))` missing at year 1. 

```



```{r misspattern, fig.align = "center",fig.cap = "\\label { } Missing data pattern.", fig.pos ='H'}

explanatory<-c("packyears","mmsebltotal","mmseyr1total")
dependent<- c("cens","tt")

mispattern<-data2 %>% 
  missing_pattern(explanatory)

```
The fig \@ref(fig:misspattern) is the visualisation of this missing pattern output, where the blue cubic means observed and red cubic means missing. From the fig \ref{fig:misspattern}, we can see that 26 patients missing with the minimental state examination score at baseline and 7 patients missing with smoking history in pack-years. The missing pattern is general and connected through the first row. The first column shows the frequency of each pattern and the number of missing entries in each pattern shows in the last column. For example, the first row means there are 241 patients in the dataset with all three observed variables. 



```{r ,eval=FALSE}

# as.data.frame(mispattern)%>%
#  rename(" "=V4)%>%
#  kable(caption='\\label{tab:mistable2}  Summary of missing data pattern.',booktabs=T,linesep="",digits=2)
```

```{r pmm}

data2$diagnosiscatbinary<-ifelse(data2$diagnosiscat=="Atypical parkinsonism",0,1)   # (,yes,no) Atypical parkinsonism=0, Parkinson's disease=1

pmm<-mice(data2[,c(4,6,9:18,25:26)],method="pmm",seed=81420,m=10,print=F)


```




```{r convergence, fig.align = "center",fig.cap = "\\label{ } Mean and standard deviation of the synthetic values plotted against iteration number for the imputed data.", fig.pos ='H'}
# pmm$visitSequence

plot(pmm)

```
From fig \@ref(fig:convergence), we can see that there are no obvious trends. Hence, the MICE algorithm has converged.


```{r densityplot1,fig.height=5, fig.width=10, fig.align = "center",fig.cap = "\\label{ } Density plot of the observed and imputed data of MMSE test score at baseline ", fig.pos ='H' }

densityplot(pmm,~ mmsebltotal,xlab=" Mini Mental State Examination test score at baseline (red line is imputed data and blue line is observed data)" )



```

From the fig \@ref(fig:densityplot1), we can see that the distribution of observed and imputed data is similar.

```{r eval=FALSE}

#densityplot2, fig.height=5, fig.width=10, fig.align = "center",fig.cap = "\\label{ } Density plot of non-zero observed (blue) and the m=10 densities per variable calculated from the imputed data (thin red lines) of packyears.", fig.pos ='H'

imp.packyears<-pmm[["imp"]][["packyears"]]


packyearsnozero<-subset(data2,packyears>0)


 plot(density(packyearsnozero$packyears),col="darkblue",main=" ", xlab = "packyears",xlim=c(-100,200),ylim=c(0.000,0.025))
for (i in 1:10) {
  lines(density(imp.packyears[,i] ),col="deeppink3")
}
 
#In this study, the `r nrow(packyearsnozero)` patients with missing “packyears” variable are all smokers, which means pack years should be larger than zero. Therefore, fig \@ref(fig:densityplot2) shows the distribution for those with non-zero "packyears". We can see that the distribution between observed and imputed data is different.

#The discrepancies may be caused by a bad imputation model, by a missing data mechanism that is not MCAR or by a combination of both. To further compare the distributions of observed and imputed data of "packyears", we use the diagnostic tool proposed by @Diagnostic2016. The idea is that under MAR the conditional distributions should be similar if the assumed model for creating multiple imputations has a good fit. 

```


```{r eval=FALSE}
#diagnostic, fig.align = "center",fig.cap = "\\label{ } Packyears against missingness probability for observed and imputed values.", fig.pos ='H'


fit<-with(pmm,glm(ici(pmm)~agebl+sex+packyears+eversmok+livesalone+charlsonbl+updrsblpart3+hybl+sebl+mmsebltotal+mmseyr1total+cens+tt,family = binomial))

# agebl+agesymptomonset+sex+packyears+eversmok+livesalone+charlsonbl+updrsblpart3+hybl+sebl+mmsebltotal+mmseyr1total+cens+tt

ps<-rep(rowMeans(sapply(fit$analyses,fitted.values)),pmm$m +1)

xyplot(pmm, packyears~ps | as.factor(.imp),
       xlab="Probability that record is incomplete",
       ylab="Packyears",
       pch=c(1,19),
       col=mdc(1:2))


#The fig \@ref(fig:diagnostic) shows that the distributions of the blue and red points are quite similar, as expected under MAR. Therefore, the imputation model has a good fit.

```





```{r mmsecat}

long<-complete(pmm, 'long', include=TRUE) #include=T means, the long contains original data and 10 imputed data.
                                          #if include=FALSE or not specified, Error in as.mids(long) : Original data not found.


N <- nrow(long)
for (i in 1:N){
  if(!is.na(long$mmsebltotal[i])){
    if (long$mmsebltotal[i] <= 23 ){
      long$mmsecat[i] = 0
    }else if (long$mmsebltotal[i] <= 27 && long$mmsebltotal[i] > 23){
      long$mmsecat[i] = 1
    }else if (long$mmsebltotal[i] <= 30 && long$mmsebltotal[i] > 27){
      long$mmsecat[i] = 2
    }
  }
}




long$mmsecat<-factor(long$mmsecat,
                        levels = c(0,1,2),
                        labels = c("<24", "24-27", "28-30"))



pmm.new<-as.mids(long)



```






**Identify the risk factors**



```{r cox models,results = 'hide'}

dd<-datadist(long)
options(datadist='dd')

modela<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+hybl+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new)  #full

modelb<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+updrsblpart3+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) #without hybl

modelc<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) #without updrsblpart3

```

```{r ,eval=FALSE}
modela
modelb
modelc

```



```{r modelselect}

AIC1<-AIC(modela)  #full
AIC2<-AIC(modelb)  #without hybl
AIC3<-AIC(modelc)  #without updrs


BIC1<-BIC(modela)  #full
BIC2<-BIC(modelb)  #without hybl
BIC3<-BIC(modelc) #without updrs


data.frame(Model=c("Full Model","Model without Hoehn & Yahr disease stage at baseline","Model without UPDRS part 3 (motor) score") , "Heuristic shrinkage"=c(0.93,0.94,0.94),'R square' = c(0.383,0.379,0.383),AIC = c(AIC1,AIC2,AIC3), BIC = c(BIC1,BIC2,BIC3))%>%
  kable(caption='Model tests and discrimination indexes.',booktabs=T,linesep="",digits=2)


```
Table \@ref(tab:modelselect) shows that the $R^2$ of these three model are the same and the heuristic shrinkage are all very close to one. Therefore, these three models are all fitted the data well and the noise is very low. And the model without UPDRS part 3 (motor) score has the lowest AIC and BIC value. Due to this study only has 140 events and to avoid over fitting, we choose the model without UPDRS part 3 (motor) score.





```{r testcox}


pmm.new_names <- paste("pmm.new",1:10,sep="")  # create names first

for (i in 1:10){
    pmm.new_df<-complete(pmm.new,i)
    assign(pmm.new_names[i], pmm.new_df)
}


test_names<- paste("test.ph",1:10,sep="")   # create names first

for (i in 1:10){
    test.ph_df<-cox.zph(coxph(Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat,data=get(paste0("pmm.new",i))),transform = "km")   # data=get(paste0("pmm.new",i))
    assign(test_names[i], test.ph_df)
}


test.ph<-unlist(test.ph10[["table"]])

test.ph<-as.data.frame(test.ph)
  
rownames(test.ph)<-c("Age","Sex","Smoking history in pack-years","Lives alone","Charlson co-morbidity index","Hoehn and Yahr Scale","Schwab and England Activities of Daily Living (ADL) Scale","Mini Mental State Examination (MMSE) test","Global") 


test.ph %>%
   kable(caption='\\label{tab:testcox} PH assumption test results.',booktabs=T,linesep="",digits=2)


```

Table \@ref(tab:testcox) shows a example of proportional hazards (PH) assumption, we can see that except for "Schwab and England Activities of Daily Living (ADL) Scale", all predictors' p-value is larger than 0.05. It indicates that only the "Schwab and England ADL Scale" effect changes significantly over time. However, the p-value of global PH assumption test is larger than 0.05. Hence there is not enough evidence to reject the null hypothesis of proportional hazards. And the p-value of global PH assumption test are all larger than 0.05 in all 10 imputed datasets. Hence, the PH assumption is valid. 



```{r,eval=FALSE}
plot(cox.zph(coxph(Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat,data=pmm.new10),transform = "km"),var="sebl")


# fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat+sebl:tt),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new)



```

```{r,eval=FALSE}

#page 545

z<-predict(modelc,type = 'terms')
z.sebl<-z[,7]
z.other<-z[,-7]
f.ia<-cph(Surv(data2$tt,data2$cens)~z.sebl*z.other)
anova(f.ia)

```








```{r martingale,fig.align = "center",fig.cap = "\\label{ } Martingale residuals versus continuous predictors.", fig.pos ='H'}

pmm.new1<-complete(pmm.new,1)
rr.0<-residuals(coxph(Surv(tt,cens)~1,data = data2),type="martingale")

pmm.new1$rr.0<-rr.0


plot1<-ggplot(pmm.new1,aes(agebl, rr.0)) + 
         geom_point(cex=1) +
         geom_smooth(method = "loess", se = T)+
         xlab("Age")+
         ylab("residuals")+
         ylim(c(-2,2))



plot2<- ggplot(pmm.new1,aes(packyears, rr.0)) + 
         geom_point(cex=1) +
         geom_smooth(method = "loess", se = T) +
         xlab("Smoking history in pack-years")+
         ylab("residuals")+
         ylim(c(-2,2))   

plot3<-ggplot(pmm.new1,aes(hybl, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        xlab("Hoehn and Yahr Scale")+
        ylab("Martingale residuals")+
        ylim(c(-2,2))


plot4<- ggplot(pmm.new1,aes(sebl, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        xlab("Schwab and England ADL Scale")+
        ylab("residuals")+
        ylim(c(-2,2))

plot5<- ggplot(pmm.new1,aes(charlsonbl, rr.0)) + 
        geom_point(cex=1) +
        geom_smooth(method = "loess", se = T)+
        xlab("Charlson co-morbidity index")+
        ylab("residuals")+
        ylim(c(-2,2))


grid.arrange(plot1,plot2,plot3,plot4,plot5, ncol=2, nrow=3)





```
From the \@ref(fig:martingale) we can see that "Smoking history in pack-years" and "Charlson co-morbidity index" are both seems to be symmetrically distributed around $\pm 1$ and with flat blue fitted line around zero. And the scatter points of "Hoehn & Yahr Scale" seems symmetrically distributed around $\pm 0.5$ with expected value of zero. For the scatter plot of "Schwab & England ADL Scale", there appears to be a slightly decreasing trend  while for "Age" the line is slightly going up. 


```{r rcs,results = 'hide'}

modelc1<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+hybl+sebl+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age as 3 knot rcs


modelc2<-fit.mult.impute((Surv(tt,cens)~agebl+sex+packyears+livesalone+charlsonbl+hybl+rcs(sebl,3)+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with sebl as 3 knot rcs


modelc3<-fit.mult.impute((Surv(tt,cens)~rcs(agebl,3)+sex+packyears+livesalone+charlsonbl+hybl+rcs(sebl,3)+mmsecat),x=T,y=T,surv=T,fitter=cph,xtrans=pmm.new) # with age and sebl as 3 knot rcs

```


```{r ,eval=FALSE}
modelc1
modelc2
modelc3

```


```{r modelselect2}
AIC.C<-AIC(modelc)    # chose one,without updrs
AIC.C1<-AIC(modelc1)  # with age as 3 knot rcs
AIC.C2<-AIC(modelc2)  # with sebl as 3 knot rcs
AIC.C3<-AIC(modelc3)  # with age and sebl as 3 knot rcs

BIC.C<-BIC(modelc)   # chose one,without updrs
BIC.C1<-BIC(modelc1)  # with age as 3 knot rcs
BIC.C2<-BIC(modelc2)  # with sebl as 3 knot rcs
BIC.C3<-BIC(modelc3)  # with age and sebl as 3 knot rcs


# restricted cubic splines (RCS)

data.frame(Model=c("Linear model","RCS model with 3 knots in age","RCS model with 3 knots in Schwab and England Activities of Daily Living (ADL) Scale","RCS model with 3 knots both in age and Schwab and England Activities of Daily Living (ADL) Scale") , 'Heuristic shrinkage'=c(0.94,0.93,0.93,0.93),'R square' = c(0.383,0.383,0.387,0.387),AIC = c(AIC.C,AIC.C1,AIC.C2,AIC.C3), BIC = c(BIC.C,BIC.C1,BIC.C2,BIC.C3))%>%
  kable(caption=' Compare model with and without Restricted cubic splines (RCS).',booktabs=T,linesep="",digits=2)

```
Table \@ref(tab:modelselect2) shows that all these four models fit the data well as their heuristic shrinkage are all very close to one. And there is only little difference between $R^2$, AIC and BIC value of models with or without RCS terms. And the linear model has the lowest BIC value and the second lowest AIC value. Therefore, since the simplest model is just as good as the models with RCS terms, we choose the linear model.




```{r coxout}

sumrcs2<-data.frame(summary(modelc))

sumrcs2.new<-sumrcs2[grep("Hazard.Ratio",rownames(sumrcs2)),]

row.names(sumrcs2.new)<-c("Age","Smoking history in pack-years","Charlson co-morbidity index","Hoehn and Yahr Scale","Schwab and England Activities of Daily Living (ADL) Scale","Sex","livesalone","Mini Mental State Examination (MMSE) test - <24:28-30 ","Mini Mental State Examination (MMSE) test - 24-27:28-30")




#sumrcs2.new%>%
#  select(Effect,Lower.0.95,Upper.0.95)%>%
#  rename("Hazard ratio"=Effect,"The 95% lower confidence limits"=Lower.0.95,"The 95% upper confidence limits"=Upper.0.95)%>%
#  kable(caption='Summary of estimated effects.',booktabs=T,linesep="",digits=2)


read.csv("Hazard ratio.csv")%>%
  rename("Characteristic"="ï..Characteristic","Hazard ratio and the 95% confidence interval"="Hazard.ratio.and.the.95..confidence.interval")%>%
  kable(caption='Summary of estimated effects.',booktabs=T,linesep="",digits=2)



  
```

Table \@ref(tab:coxout) shows that patients' with with mild dementia (less than 24 score in MMSE test) have the highest hazard ratio (HR=3.12) compare to patients with normal cognitive function (MMSE test score larger than 28). The decline in cognitive function (MMSE test score from 24 to 27) increase the risk of entering the nursing home (HR=1.85). The second highest risk for patient's need of long-term care is ageing (HR=2.01). The slightly significant risk factor is smoking habit as its 95% lower confidence interval just above one and the hazard ratio is closed to one (HR=1.18). However, patient's sex and whether lives alone are both don't have significant effect on the chance of entering into institutionalisation as their 95% hazard ratio confidence interval both include one. Woman have smoking habit and lives alone tend to have higher need of long-term care. For the patients' annual clinical examinations of disease stage (Hoehn and Yahr Scale), disability (Schwab and England Activities of Daily Living (ADL) Scale), and comorbid conditions (Charlson co-morbidity index) are all not statistical significant as their 95% hazard ratio confidence interval all include one. The point estimate hazard ratio indicates that less independence patients (lower percentages in Schwab and England ADL Scale) and more serious disease stage patients have higher chance of being admitted into nursing home. However, patients with more serious comorbid conditions have less chance entering into long term care facilities. 

```{r coxoutplot,out.width="70%",fig.align = "center",fig.cap = "\\label{ } Hazard ratios and confidence bars for risk factors in model.", fig.pos ='H'}


text<-"Notes:\nagebl is age; packyears is smoking history in pack-years;\ncharlsonbl is Charlson co-morbidity index; hybl is Hoehn and Yahr Scale;\nsebl is Schwab and England Activities of Daily Living (ADL) Scale;\nmmsecat is Mini Mental State Examination (MMSE) test."
par(mar = c(8,5,3,3))
plot(summary (modelc),main= '',log = T)
mtext(text, side = 1, line = 5, cex = 0.8, adj = 0) 
```

The \@ref(fig:coxoutplot) shows the estimated hazard ratios of patients' default setting of predictors at baseline visit. Every 10 years older for patients, the risk of entering the nursing home will be double. The risk of institutionalisation for patients who smoke 23 packs in a year only slight higher than patients without smoking habit. Every two Charlson co-morbidity index higher for patients, the need of institutionalisation is decrease around 20%. Patients with one more stage of Hoehn and Yahr Scale or 30 lower percentage score in Schwab and England ADL Scale will increase around 25% risk of institutionalisation. 


```{r nomogram, eval=FALSE}



surv <- Survival(modelc)
surv5 <- function(x) surv(5*365,lp=x)
surv8 <- function(x) surv(8*365,lp=x)


plot(nomogram(modelc, 
              fun=list(surv5, surv8), 
              funlabel=c('5-year survival', '8-year survival')))

```



```{r, eval=FALSE}


img1_path<-"nomogram.png"
img1<-readPNG(img1_path,native = T,info = T)
# attr(img1,"info")
include_graphics(img1_path)
```





## Discussion 


This study found that patients with atypical parkinsonism have 5.3 times higher incidence rate of entering the long-term care facilities and 4.7 times shorter median survival time to institutionalisation than patients with Parkinson’s disease (PD). The highest hazard ratio risk factor is mild dementia (HR=3.1), follows with ageing (HR=2.0) and cognitive function declines (HR=1.2), and these three risk factors are three statistical significance as their 95% confidence interval don't include one. Female patients who live alone and have smoking habit are more likely need long-term care. And patients with more serious disability and disease stage have higher chance of admitting into nursing home. However, patients with more serious comorbid conditions tend to have less chance of entering institutionalisation. 


The 













# Reference


